### 2014-06-11 paper.tex
refining writing

% This version uses the latex2e styles, not the very ancient 2.09 stuff.
\documentclass{sig-alternate-10pt}
\usepackage{endnotes,url}
%\usepackage{sig-alternate-10pt,endnotes}
\usepackage{epsfig}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{verbatim}
% Replaced by subfig package
% \usepackage{subfigure}
\usepackage{subfig}
\usepackage{epstopdf}
\usepackage{cite}
\usepackage{array}
\usepackage{multirow}
\usepackage{listings}

%\pdfpagewidth 8.5in
%\pdfpageheight 11in
%\baselineskip=12.5pt
%\setlength{\pdfpagewidth}{8.5in}
%\setlength{\pdfpageheight}{11in}
%\usepackage{pdfdraftcopy}

% To import classical math notation used
\input{mathnotation.tex}
% Package for date and time presentation
\usepackage{datetime}
\newcommand{\datation}{\today, \xxivtime}

\def\full{0}        % set 1 for a full tech report version
                    % set 0 for submission version
\def\shownotes{1}   % set 1 for version with author notes
                    % set 0 for no notes
\def\anon{0}        % set 1 to anonymize
                    % set 0 for acks and author names

%%%%%%%  Author Notes %%%%%%%
%
\ifnum\shownotes=1
\newcommand{\authnote}[2]{{ $\ll$\textsf{\footnotesize #1 notes: #2}$\gg$}}
\else
\newcommand{\authnote}[2]{}
\fi
\newcommand{\Ynote}[1]{{\authnote{Yan}{#1}}}
\newcommand{\Anote}[1]{{\authnote{Andrius}{#1}}}
\newcommand{\Nnote}[1]{{\authnote{Narseo}{#1}}}
\newcommand{\Vnote}[1]{{\authnote{Vijay}{#1}}}
\newcommand{\Dnote}[1]{{\authnote{Dina}{#1}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\namedref}[2]{#1~\ref{#2}}
\newcommand{\tableref}[1]{\namedref{Table}{#1}}
\newcommand{\sectionref}[1]{\namedref{Section}{#1}}
\newcommand{\appendixref}[1]{\namedref{Appendix}{#1}}
\newcommand{\theoremref}[1]{\namedref{Theorem}{#1}}
\newcommand{\remarkref}[1]{\namedref{Remark}{#1}}
\newcommand{\definitionref}[1]{\namedref{Definition}{#1}}
\newcommand{\figureref}[1]{\namedref{Figure}{#1}}
\newcommand{\lemmaref}[1]{\namedref{Lemma}{#1}}
\newcommand{\claimref}[1]{\namedref{Claim}{#1}}
\newcommand{\propositionref}[1]{\namedref{Proposition}{#1}}
\newcommand{\constructionref}[1]{\namedref{Construction}{#1}}
\newcommand{\corollaryref}[1]{\namedref{Corollary}{#1}}
\newcommand{\equationref}[1]{\namedref{Equation}{#1}}
%
\newtheorem{theorem}{Theorem}[section]
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{obs}[theorem]{Observation}
%

\providecommand{\vs}{vs. }
\providecommand{\ie}{\emph{i.e.,} }
\providecommand{\eg}{\emph{e.g.,} }
\providecommand{\cf}{\emph{cf.,} }
\providecommand{\resp}{\emph{resp.,} }
\providecommand{\etal}{\emph{et al.}}   %Removed trailing space here; usually want non-breaking space with following reference
\providecommand{\etc}{\emph{etc.}}      % No trailing space here either
\providecommand{\mypara}[1]{\smallskip\noindent\emph{#1} }
\providecommand{\myparab}[1]{\smallskip\noindent\textbf{#1} }
\providecommand{\myparasc}[1]{\smallskip\noindent\textsc{#1} }
\providecommand{\para}{\smallskip\noindent}

\newtheorem{axiom}{{\bf  Axiom}}
\newtheorem{defin}{{\bf  Definition}}
\newtheorem{proposition}{Proposition}

\usepackage{enumitem}
\setlist{nolistsep}

%Model parameters

\usepackage[breaklinks=true]{hyperref}
\frenchspacing
\begin{document}

%don't want date printed
\date{}

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{PhoneLets: offloading the phone off your phone for energy, cost and network load optimisation}
\ifnum\anon=1
\author{[Paper: \#XXX]}% \hspace{0.2cm} \ampmtime ]}
\else
\numberofauthors{2}
\author{
% \alignauthor Andrius Aucinas\\
% \affaddr{University of Cambridge} 
% \and
% \alignauthor Jon Crowcroft\\
% \affaddr{University of Cambridge}
}
\fi
%for single author (just remove % characters)

% end author
\maketitle
% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
%\thispagestyle{empty}
\begin{abstract}
This demo presents \emph{PhoneLets}, a system to provide offloading of phone functionality of mobile cellular devices to a neighboring, potentially powered one. Phone offloading happens through providing access to cellular network subscriber credentials over a short-range wireless link and selecting the device that uses them to register to a particular network. As a result, we can use a single identity (SIM card) across multiple clients without increasing their connectivity costs, the load for the network handling large numbers of other subscribers and we are able to optimize mobile clients' energy consumption by leveraging powered \emph{PhoneLets} or using them to share the energy consumption load across distributed nodes.
\end{abstract}

\section{What PhoneLets are}
\label{section:intro}

Not only the modern smartphone has more in common with a computer than a phone, but connectivity to the Internet as well as the increasing number of surrounding devices (the \emph{Internet of Things}) is tipping the balance even further. Cellular networks, however, are just as important as they provide almost ubiquitous Internet connectivity around the world. There are multiple problems with cellular connectivity.

Firstly, it is relatively expensive for the users~\cite{Anonymous:2013ut} and typically the end user needs to obtain a separate subscription for each of his devices connected to the cellular network, e.g. smartphone and tabled as well as broadband connectivity at home and at work, each with its underutilized caps and subscription fees.

Secondly, the price is largely driven by the high infrastructure costs which increase with the growing number of network users - the increasing number of connected clients periodically communicating data also increase cellular networks loads, in the extreme cases causing \emph{signaling storms}~\cite{Yang:wh} that can bring the entire network down.

Finally, energy consumption of always connected clients is high compared to short-range radio technologies. We have analyzed in the past how power-hungry such communication can be from the mobile device perspective~\cite{Aucinas:2013uk} and without careful optimization of such communication patterns the cost may be too high for many types of devices.

These factors all motivate the need for \emph{PhoneLets}, a wireless system that enables offloading of cellular connectivity off a device to share a subscription between devices, balance energy consumption of a group of devices or simply replace the cellular link with a short-range link to a powered device.

\emph{PhoneLets} enable the functionality by exploiting remote SIM functionality of certain hardware, whereby a cellular module accepts to exchange commands with a SIM/UICC module over a separate communication link rather than direct, physical connection. Building our prototype on the existing technology allows us to achieve more advanced functionality without breaking encryption of participating entities or doing unlawful communication interception. Instead, \emph{PhoneLets} infrastructure establishes and manages secure communication channels between participating entities and allows a device to connect to a cellular network without a physical SIM module being present.

\section{Strawman Design}

\begin{figure}
\centering
\includegraphics[width=0.95\columnwidth]{figs/arch}
\caption{Architecture}
\label{fig:arch}
\end{figure}

The overall concept and system architecture is shown in Fig.~\ref{fig:arch}. There are three types of entities: mobile server - a mobile device which can share its SIM with other clients, SIM server which may host and provide more sets of credentials over local or wide-area networks and a mobile client which can use either of the methods.

Atypically for cellular networking technologies, certain parts are relatively loosely coupled as the baseband does not require physical presence of a SIM  or relays communication to its SIM. Furthermore the SIM is a smartcard which can be interfaced with a standard smartcard reader, allowing for a relatively simple server design. Our work for this demo focused primarily on the client and SIM server design, and we are using off-the-shelf hardware as the mobile server to show compatibility with the specification.

\begin{figure}[t!]
\centering
\includegraphics[width=0.95\columnwidth]{figs/client}
\caption{Client prototype}
\label{fig:client}
\end{figure}

Our demo mobile client (Fig.~\ref{fig:client}) consists of a Telit UC864 cellular module connected to a raspberryPI as a controller, which also uses WiFi and BlueTooth dongles for local wireless connectivity. The controller runs a modified version of the Ofono open telephony stack\footnote{\url{https://ofono.org/}} and manages communication between the cellular module and multiple backends.

In addition to an unmodified Galaxy S4 running as the mobile client, we have also implemented a server that does not require a cellular module and instead interfaces with SIM cards through a smartcard reader. Once a connection is established to either backend (over BlueTooth to the mobile server or an encrypted connection to the SIM server), the cellular module communicates with the remote card using the standard Application Protocol Data Unit (APDU) messages (Fig.~\ref{fig:sequence}) and registers with the network. Importantly, in this case we replace the BlueTooth communication channel with an arguably more secure option in that we use certificate-based authentication of both client and server and use strong SSL/TLS encryption for all data exchange.

An important point to make is that we do not clone the SIM card and by design the system disables connectivity for one device before enabling it for another one to prevent multiple devices sharing the same credentials simultaneously as this would very likely be unsupported by the network even if it were not considered to be an attempt to illegally clone a card. Therefore, one device is connected to the cellular network and may share (\emph{tether}) the connectivity to other devices over another channel. One cellular module is also only capable of being registered with one network, therefore enforcing a strict one-to-one mapping.

One part of the network's security is that the network authenticates the subscriber for any major operation (e.g. placing or receiving phonecalls, establishing PDP context) as well as periodically checks for subscriber's presence, therefore the link between the client and the corresponding server must be always maintained, simplifying security requirements of the communication channel.

\begin{figure}[t!]
\centering
\includegraphics[width=0.95\columnwidth]{figs/sequence}
\caption{Communication sequence between client, server and remote SIM card}
\label{fig:sequence}
\end{figure}

\section{Potential for energy efficiency}

A major motivation for using \emph{PhoneLets} is reduced mobile client energy consumption. In this section we provide energy consumption of a few example scenarios using our early prototype. Since the client itself is not yet energy-optimized, i.e. does not use the full energy saving capabilities of the radio module and the controlling RaspberryPI is not meant for mobile uses, we focused on energy comparison of the mobile SIM server.

\begin{table}[t]
{
\small
\begin{tabular}{| l | c |}
\hline
  \textbf{Use case}         & \textbf{Avg. energy consumption (mA)}  \\ \hline
  Idle - airplane mode      & 5  \\ \hline
  Only BlueTooth on         & 6  \\ \hline
  Idle - only WiFi          & 8  \\ \hline
  Idle - 3G on              & 10 \\ \hline
  Apps background           & 22.4 - 54.5  \\ \hline
  WiFi Apps background      & ??    \\ \hline
  WiFi frequent messages    & YY    \\ \hline
  BT SIM access             & 28  \\ \hline
  WiFi tethering (idle)     & 80  \\ \hline
\end{tabular}
}
\caption{Energy consumption of the different usage scenarios of a mobile device.}
\label{tab:energy}
\end{table}

We measured energy consumption of a Samsung Galaxy S4 phone using the Power Monitor\footnote{Monsoon Power Monitor \url{http://www.msoon.com/LabEquipment/PowerMonitor/}}. The different values are presented in Table.~\ref{tab:energy}. Energy consumption of an idle device with applications in background (\emph{Apps background} considers a device with Google Apps, Skype, Facebook and Viber apps installed and running in background and is from our previous work~\cite{Aucinas:2013uk}.

With the default implementation of Remote SIM Access over BlueTooth, the device consumes three times the amount of energy of an idle device connected to a 3G network, which is a high energy cost to pay, however the amount is similar or even less than that of a device with multiple applications running in background but otherwise idle. Furthermore, energy consumption of WiFi tethering even when a client connected to it is idle is significantly higher than that of sharing connectivity by offloading the phone functionality.

These are early results, and we have not yet been able to replace the BlueTooth channel for an Android device. Nevertheless, the results point to a potential use-case where a powered \emph{PhoneLet} would be registered to the network on behalf of a WiFi-connected client with a considerable energy benefit: client connected over WiFi and exchanging messages of the same size and frequency as required by the network consumes YYmA of energy on average. The benefit is further improved when network access is shared across multiple such devices and SIM access load is balanced across them.

\section{Demo: wireless SIM sharing}

To demonstrate the working prototype of the system, we show our client \emph{PhoneLet} (Fig.~\ref{fig:client}), communicating either with an off-the shelf mobile device, or a SIM server running on a raspberryPI to share their
credentials to connect to their specific networks. We do so in three steps:

\begin{enumerate}
    \item Make a phone call to the mobile server (phone) and show that it is ringing
    \item Start sharing its SIM card with the client and show that the client is registered to the network with the same ID by making a phone call to it
    \item Start using SIM card in the SIM server and show that the client is registered to the network with a different ID by making a call to the other number.
\end{enumerate}

In the three number of steps we show our prototype's ability to offload phone functionality off a phone - share subscriber credentials between multiple clients and select which one will be connected to the network.

\section{Discussion}

If we only consider Internet connectivity, one could claim that \emph{PhoneLets} can be replaced by simple tethering as supported by all modern smartphones, however it would not take into account battery consumption, resource savings from sharing a single connection or even the lower energy consumption of sharing the SIM card than sharing Internet connectivity as discussed above.

As much as the telephone number is important, multiple providers have offered a service in the past whereby multiple SIM cards would be associated with the number and multiple handsets would ring on an incoming call\footnote{Google have also offered such service with Google Voice \url{http://www.google.co.uk/googlevoice/about.html\#tab=oneNumber}}, however it does not solve the problem of an increasing number of cellular clients and billing depends on the exact agreement with the service provider. A cellular network in UK has even offered a service to make phonecalls from tablets and laptops over WiFi to its users using a special app\footnote{O2 TU Go \url{http://www.o2.co.uk/apps/tu-go}}, decoupling number from the phone even further.

Reprogrammable SIM cards~\cite{OFcom:2012tx} and more recently embedded SIM~\cite{Association:2013ub} have been proposed as a solution to the provisioning problem for M2M communication. Essentially, it still uses physical secure element to store credentials, but provides mechanisms to manage them over the air, primarily for long lifetime applications where replacing SIM cards is difficult. The architecture, unsurprisingly, allows operators to retain full control over the credentials, including when new ones can be installed or old ones removed, necessitating centralized management and reducing flexibility on the client side.

Finally, security is a major concern and the reason why secure elements are used by operators in the first place despite multiple proposals to move towards software-based SIMs. In our proposal, however, we still maintain the separate secure element and only replace the interface to it with an arguably secure alternative widely used in industry. Furthermore, past proposals of remote SIM usage have already included mobile payment solutions~\cite{subramanian2003sim}, which are even more risky than communication alone. Detailed security analysis is subject to future work, but we believe that any additional risk created by \emph{PhoneLets} is marginal.

\section{Summary}

A peculiar feature of the smartphone has been the decreasing importance of it as a phone, instead becoming an Internet device. Subscription to cellular networks, however, becomes expensive and energy-consuming for multiple devices. \emph{PhoneLets} is a system that allows for offloading of phone functionality of a cellular device, instead sharing a single subscription and therefore a single link amongst multiple devices in order to save costs and energy, decreasing cellular network infrastructure load at the same time.

More information and demonstration video is available via \url{phonelets.smart-e.org}.

%We highly encourage that each proposal includes a video clip showcasing the demo. The video clip, no more than 60 seconds, should give a good idea of what the demo is about and how it would look like. If you do include a video clip, it will help the committee better understand and evaluate your proposal. If your demo is accepted, the video will likely be shown on the conference website and during the conference to draw people to your booth. Videos can be submitted by including a link to the video (e.g., hosted at YouTube).

\section*{Resource Requirements}

(1) \emph{Desk/table of approximately 2m$^2$} to accommodate the demo. (2) \emph{Two standard UK power sockets}. (3) \emph{Two cellular network subscriptions (SIM cards)}, but are to be sources by the presenters. Setup time - up to an hour for connecting the demo up and verifying operation. Other equipment includes a laptop, two smartphones and the custom-build hardware described above.

% Equipment to be used for the demo.
% Space needed and setup time required.
% Additional facilities needed including power and Internet/wireless access.

%Only show things we really cite :)
% \nocite{*}
{\footnotesize 
\bibliographystyle{abbrv}
%\footnotesize
\bibliography{references}
}

%\appendix 
%\input{data}

\end{document}

### 2014-06-13 paper.tex
a few tweaks, wifi and 3g numbers adjusted for gs4

@@
 %make title bold and 14 pt font (Latex default is non-bold, 16 pt)
-\title{PhoneLets: offloading the phone off your phone for energy, cost and network load optimisation}
+\title{PhoneLets: offloading the phone off your phone for energy, cost and network load optimization}
 \ifnum\anon=1
@@
 \author{
-% \alignauthor Andrius Aucinas\\
-% \affaddr{University of Cambridge} 
-% \and
-% \alignauthor Jon Crowcroft\\
-% \affaddr{University of Cambridge}
+\alignauthor Andrius Aucinas\\
+\affaddr{University of Cambridge} 
+\and
+\alignauthor Jon Crowcroft\\
+\affaddr{University of Cambridge}
 }
@@
 
-\section{Potential for energy efficiency}
+\section{Energy saving opportunities}
 
@@
   Idle - airplane mode      & 5     \\ \hline
-  Only BlueTooth on         & 6  \\ \hline
-  Idle - only WiFi          & 8  \\ \hline
-  Idle - 3G on              & 10 \\ \hline
-  Apps background           & 22.4 - 54.5  \\ \hline
-  WiFi Apps background      & ??    \\ \hline
+  Ifle - BlueTooth          & 6     \\ \hline
+  Idle - WiFi               & 8     \\ \hline
+  Idle - 3G                 & 10    \\ \hline
+  Apps background - WiFi    & 27    \\ \hline
+  BT SIM access             & 28    \\ \hline
   WiFi frequent messages    & YY    \\ \hline
-  BT SIM access             & 28  \\ \hline
+  Apps background - 3G      & 40    \\ \hline
   WiFi tethering (idle)     & 80    \\ \hline
@@
 The different values are presented in Table.~\ref{tab:energy}.
-Energy consumption of an idle device with applications in background (\emph{Apps background} considers a device with Google Apps, Skype, Facebook and Viber apps installed and running in background and is from our previous work~\cite{Aucinas:2013uk}.
+Energy consumption of an idle device with applications in background (\emph{Apps background}) considers a device with Google Apps, and Facebook installed and running in background and is consistent with our earlier observations~\cite{Aucinas:2013uk}.
+We have only run the test on a single WiFi network, therefore we expect the figure to vary, especially because our used network is a public one with some background traffic affecting energy consumption of the device.
 
@@
 
-(1) \emph{Desk/table of approximately 2m$^2$} to accommodate the demo.
+(1) \emph{Desk/table of approximately 1.5m, standard-width} to accommodate the demo.
 (2) \emph{Two standard UK power sockets}.
-(3) \emph{Two cellular network subscriptions (SIM cards)}, but are to be sources by the presenters.
+(3) \emph{Two cellular network subscriptions (SIM cards)}, but are to be sources by the presenters, (4) an HDMI display during setup.
 Setup time - up to an hour for connecting the demo up and verifying operation.

### 2014-06-16 rpi-sim/dbus-client.py
bootstrapping mobicom demo repo, working python server over zmq

@@ -1,0 +1,80 @@
+#!/usr/bin/env python
+
+import sys
+from traceback import print_exc
+from smartcard.util import toHexString
+import numpy as np
+
+import dbus
+
+HWHEADER = [0x05, 0x01, 0x00, 0x00, 0x04, 0x00, 0x00]
+commands = [
+    [00, 01, 00, 00, 00, 00, 00, 02, 01, 0x2C, 00, 00],
+    [07, 00, 00, 00],
+    HWHEADER+[7]+[0x00, 0xA4, 0x00, 0x04, 0x02, 0x3F, 0x00]+[0],
+    HWHEADER+[7]+[0x00, 0xA4,],
+    [0x08, 0x04, 0x02, 0x2F, 0xE2]+[0],
+    HWHEADER+[5]+[0x00, 0xB0, 0x00, 0x00, 0x0A, 0, 0]+[0],
+    HWHEADER+[25]+[0x80, 0x10, 0x00, 0x00, 0x14, 0xFF, 0x9F,],
+    [0xFF, 0xFF, 0x7F, 0x9F, 0x00, 0xDF, 0xFF, 0x03, 0x07, 0x1F, 0x00, 0x08, 0x0C, 0x06, 0x00, 0xEB, 0x0E, 0x03]+[0, 0, 0],
+    # [0x80, 0x12, 0x00, 0x00, 0x0F],
+    # [0x00, 0xA4, 0x08, 0x04, 0x02, 0x2F, 0x00,],
+    # [0x80, 0x14, 0x00, 0x00, 0x0C, 0x81, 0x03, 0x01, 0x05, 0x00, 0x02, 0x02, 0x82, 0x81, 0x03, 0x01, 0x00,],
+    # [0x00, 0xA4, 0x08, 0x04, 0x02, 0x2F, 0x06,],
+    # [0x00, 0xB2, 0x03, 0x04, 0xC3,],
+    # [0x00, 0xA4, 0x08, 0x0C, 0x02, 0x2F, 0x00,],
+    # [0x00, 0xB2, 0x01, 0x04, 0x26,],
+    # [0x00, 0xA4, 0x00, 0x0C, 0x02, 0x3F, 0x00,],
[…]

### 2014-06-16 rpi-sim/dbus-service.py

@@ -1,0 +1,317 @@
+#!/usr/bin/env python
+
+import threading
+import Queue
+import math
+
+import gobject
+import dbus
+import dbus.service
+import dbus.mainloop.glib
+from smartcard.CardType import AnyCardType
+from smartcard.CardRequest import CardRequest
+from smartcard.CardConnectionObserver import CardConnectionObserver
+from smartcard.CardConnectionObserver import ConsoleCardConnectionObserver
+from smartcard.util import toHexString
+
+from smartcard.sw.ErrorCheckingChain import ErrorCheckingChain
+from smartcard.sw.ISO7816_4ErrorChecker import ISO7816_4ErrorChecker
+from smartcard.sw.ISO7816_8ErrorChecker import ISO7816_8ErrorChecker
+from smartcard.sw.SWExceptions import SWException, WarningProcessingException
+
+class RSAPMessageProtocol:
+    # Message is of a format:
+    # |  1  |  1   |    2   |        varies          |
+    #  msgID #param reserved     payload 
+    #
+    # Payload (parameter) looks like
+    # |  1  |  1  |    2    |     varies      | 0-3 |  
+    #  param  rsv. paramLen    parameter value  padding
+    #   ID
+    # The length of each parameter will be a multiple of 4 bytes - use padding to reach that
+    def __init__(self):
+        self.messageID = {
+            0x00: 'CONNECT_REQ',
+            0x01: 'CONNECT_RESP',
+            0x02: 'DISCONNECT_REQ',
+            0x03: 'DISCONNECT_RESP',
+            0x04: 'DISCONNECT_IND',
+            0x05: 'TRANSFER_APDU_REQ',
+            0x06: 'TRANSFER_APDU_RESP',
+            0x07: 'TRANSFER_ATR_REQ',
+            0x08: 'TRANSFER_ATR_RESP',
+            0x09: 'POWER_SIM_OFF_REQ',
+            0x0A: 'POWER_SIM_OFF_RESP',
+            0x0B: 'POWER_SIM_ON_REQ',
+            0x0C: 'POWER_SIM_ON_RESP',
+            0x0D: 'RESET_SIM_REQ',
+            0x0E: 'RESET_SIM_RESP',
+            0x0F: 'TRANSFER_CARD_READER_STATUS_REQ',
+            0x10: 'TRANSFER_CARD_READER_STATUS_RESP',
+            0x11: 'STATUS_IND',
[…]

### 2014-06-16 rpi-sim/dbusrelaynet.py

@@ -1,0 +1,120 @@
+#!/usr/bin/env python
+
+import threading
+import Queue
+import math
+
+import gobject
+import dbus
+import dbus.service
+import dbus.mainloop.glib
+
+import zmq
+import zmq.auth
+import sys
+import logging
+import os
+from zmq.auth.thread import ThreadAuthenticator
+from smartcard.util import toHexString
+
+class DemoException(dbus.DBusException):
+    _dbus_error_name = 'org.smart_e.DemoException'
+
+class Server(dbus.service.Object):
+    def __init__(self):
+        bus_name = dbus.service.BusName("org.smart_e.RSAP", bus=dbus.SystemBus())
+        dbus.service.Object.__init__(self, bus_name, '/RSAPServer')
+        self.buffer = []
+        self.state = "CONNECT_REQ"
+        self.reqQueue = Queue.Queue()
+        self.respQueue = Queue.Queue()
+
+        # These direcotries are generated by the generate_certificates script
+        base_dir = os.path.dirname(__file__)
+        keys_dir = os.path.join(base_dir, 'certificates')
+        public_keys_dir = os.path.join(base_dir, 'public_keys')
+        secret_keys_dir = os.path.join(base_dir, 'private_keys')
+
+        if not (os.path.exists(keys_dir) and os.path.exists(keys_dir) and os.path.exists(keys_dir)):
+            logging.critical("Certificates are missing - run generate_certificates.py script first")
+            sys.exit(1)
+
+        ctx = zmq.Context().instance()
+
+        # Start an authenticator for this context.
+        auth = ThreadAuthenticator(ctx)
+        auth.start()
+        # auth.allow('127.0.0.1')
+        # Tell authenticator to use the certificate in a directory
+        auth.configure_curve(domain='*', location=public_keys_dir)
+
+        client = ctx.socket(zmq.REQ)
+
+        # We need two certificates, one for the client and one for
+        # the server. The client must know the server's public key
+        # to make a CURVE connection.
+        client_secret_file = os.path.join(secret_keys_dir, "client.key_secret")
+        client_public, client_secret = zmq.auth.load_certificate(client_secret_file)
+        client.curve_secretkey = client_secret
+        client.curve_publickey = client_public
+
+        server_public_file = os.path.join(public_keys_dir, "server.key")
+        server_public, _ = zmq.auth.load_certificate(server_public_file)
+        # The client must know the server's public key to make a CURVE connection.
[…]

### 2014-06-16 rpi-sim/simclient.py

@@ -1,0 +1,121 @@
+#!/usr/bin/env python
+
+import sys
+from traceback import print_exc
+from smartcard.util import toHexString
+import numpy as np
+
+import zmq
+import zmq.auth
+import sys
+import logging
+import os
+from zmq.auth.thread import ThreadAuthenticator
+
+HWHEADER = [0x05, 0x01, 0x00, 0x00, 0x04, 0x00, 0x00]
+commands = [
+    [00, 01, 00, 00, 00, 00, 00, 02, 01, 0x2C, 00, 00],
+    [07, 00, 00, 00],
+    HWHEADER+[7]+[0x00, 0xA4, 0x00, 0x04, 0x02, 0x3F, 0x00]+[0],
+    HWHEADER+[7]+[0x00, 0xA4,],
+    [0x08, 0x04, 0x02, 0x2F, 0xE2]+[0],
+    HWHEADER+[5]+[0x00, 0xB0, 0x00, 0x00, 0x0A, 0, 0]+[0],
+    HWHEADER+[25]+[0x80, 0x10, 0x00, 0x00, 0x14, 0xFF, 0x9F,],
+    [0xFF, 0xFF, 0x7F, 0x9F, 0x00, 0xDF, 0xFF, 0x03, 0x07, 0x1F, 0x00, 0x08, 0x0C, 0x06, 0x00, 0xEB, 0x0E, 0x03]+[0, 0, 0],
+    # [0x80, 0x12, 0x00, 0x00, 0x0F],
+    # [0x00, 0xA4, 0x08, 0x04, 0x02, 0x2F, 0x00,],
+    # [0x80, 0x14, 0x00, 0x00, 0x0C, 0x81, 0x03, 0x01, 0x05, 0x00, 0x02, 0x02, 0x82, 0x81, 0x03, 0x01, 0x00,],
+    # [0x00, 0xA4, 0x08, 0x04, 0x02, 0x2F, 0x06,],
+    # [0x00, 0xB2, 0x03, 0x04, 0xC3,],
+    # [0x00, 0xA4, 0x08, 0x0C, 0x02, 0x2F, 0x00,],
+    # [0x00, 0xB2, 0x01, 0x04, 0x26,],
+    # [0x00, 0xA4, 0x00, 0x0C, 0x02, 0x3F, 0x00,],
[…]

### 2014-06-16 rpi-sim/simserver.py

@@ -1,0 +1,334 @@
+#!/usr/bin/env python
+
+import threading
+import Queue
+import math
+import numpy as np
+
+from smartcard.CardType import AnyCardType
+from smartcard.CardRequest import CardRequest
+from smartcard.CardConnectionObserver import CardConnectionObserver
+from smartcard.CardConnectionObserver import ConsoleCardConnectionObserver
+from smartcard.util import toHexString
+
+from smartcard.sw.ErrorCheckingChain import ErrorCheckingChain
+from smartcard.sw.ISO7816_4ErrorChecker import ISO7816_4ErrorChecker
+from smartcard.sw.ISO7816_8ErrorChecker import ISO7816_8ErrorChecker
+from smartcard.sw.SWExceptions import SWException, WarningProcessingException
+
+import zmq
+import zmq.auth
+import sys
+import logging
+import os
+from zmq.auth.thread import ThreadAuthenticator
+
+class RSAPMessageProtocol:
+    # Message is of a format:
+    # |  1  |  1   |    2   |        varies          |
+    #  msgID #param reserved     payload 
+    #
+    # Payload (parameter) looks like
+    # |  1  |  1  |    2    |     varies      | 0-3 |  
+    #  param  rsv. paramLen    parameter value  padding
+    #   ID
+    # The length of each parameter will be a multiple of 4 bytes - use padding to reach that
+    def __init__(self):
+        self.messageID = {
+            0x00: 'CONNECT_REQ',
+            0x01: 'CONNECT_RESP',
+            0x02: 'DISCONNECT_REQ',
+            0x03: 'DISCONNECT_RESP',
+            0x04: 'DISCONNECT_IND',
+            0x05: 'TRANSFER_APDU_REQ',
+            0x06: 'TRANSFER_APDU_RESP',
+            0x07: 'TRANSFER_ATR_REQ',
+            0x08: 'TRANSFER_ATR_RESP',
+            0x09: 'POWER_SIM_OFF_REQ',
+            0x0A: 'POWER_SIM_OFF_RESP',
+            0x0B: 'POWER_SIM_ON_REQ',
+            0x0C: 'POWER_SIM_ON_RESP',
+            0x0D: 'RESET_SIM_REQ',
+            0x0E: 'RESET_SIM_RESP',
+            0x0F: 'TRANSFER_CARD_READER_STATUS_REQ',
[…]

### 2014-06-16 ofono-bt/plugins/telit.c
applied patch to make telit module work with remote sim over bluetooth + debugging info

@@ -83,6 +83,16 @@ static void telit_debug(const char *str, void *user_data)
 	ofono_info("%s%s", prefix, str);
 }
 
+static void hex_print(const char *buf, const gsize bytes_read) {
+    char buf_str[300 * 3 + 1] = { 0 };
+    char *endBuf = buf_str;
+    int i;
+    for (i = 0; i < bytes_read; i++) {
+        endBuf += sprintf(endBuf, "%02X ", (unsigned char)buf[i]);
+    }
+    DBG("read %zu bytes: %s", bytes_read, buf_str);
+}
+
 static void sap_close_io(struct ofono_modem *modem)
 {
 	struct telit_data *data = ofono_modem_get_data(modem);
@@ -129,9 +139,17 @@ static gboolean bt_event_cb(GIOChannel *bt_io, GIOCondition condition,
 		status = g_io_channel_read_chars(bt_io, buf, 300,
 							&bytes_read, NULL);
 
-		if (bytes_read > 0)
-			g_io_channel_write_chars(data->hw_io, buf,
+		if (bytes_read > 0) {
+			GIOStatus st;
+			st = g_io_channel_write_chars(data->hw_io, buf,
 					bytes_read, &bytes_written, NULL);
+			DBG("BT event <--- (read %d, wrote %d bytes to hw_io)", (int)bytes_read, (int)bytes_written);
+       		hex_print(buf, bytes_read);
+       		if (st != G_IO_STATUS_NORMAL && st != G_IO_STATUS_AGAIN) {
+       			DBG("Bad status");
+				return FALSE;
+       		}
+		}
 
 		if (status != G_IO_STATUS_NORMAL && status != G_IO_STATUS_AGAIN)
 			return FALSE;
@@ -166,9 +184,12 @@ static gboolean hw_event_cb(GIOChannel *hw_io, GIOCondition condition,
 		status = g_io_channel_read_chars(hw_io, buf, 300,
 							&bytes_read, NULL);
 
-		if (bytes_read > 0)
+		if (bytes_read > 0) {
 			g_io_channel_write_chars(data->bt_io, buf,
 					bytes_read, &bytes_written, NULL);
+			DBG("HW event --->");
+			hex_print(buf, bytes_read);
+		}
 
 		if (status != G_IO_STATUS_NORMAL && status != G_IO_STATUS_AGAIN)
 			return FALSE;
[…]

### 2014-06-16 ofono-wifi/drivers/atmodem/network-registration.c
copied ofono into a separate tree for the wifi version

@@ -1,0 +1,2040 @@
+/*
+ *
+ *  oFono - Open Source Telephony
+ *
+ *  Copyright (C) 2008-2011  Intel Corporation. All rights reserved.
+ *  Copyright (C) 2010  ST-Ericsson AB.
+ *
+ *  This program is free software; you can redistribute it and/or modify
+ *  it under the terms of the GNU General Public License version 2 as
+ *  published by the Free Software Foundation.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU General Public License for more details.
+ *
+ *  You should have received a copy of the GNU General Public License
+ *  along with this program; if not, write to the Free Software
+ *  Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA  02110-1301  USA
+ *
+ */
+
+#ifdef HAVE_CONFIG_H
+#include <config.h>
+#endif
+
+#define _GNU_SOURCE
+#include <string.h>
+#include <stdlib.h>
+#include <stdio.h>
+
+#include <glib.h>
+
+#include <ofono/log.h>
+#include <ofono/modem.h>
+#include <ofono/netreg.h>
+
+#include "gatchat.h"
+#include "gatresult.h"
+
+#include "common.h"
+#include "atmodem.h"
+#include "vendor.h"
+
+static const char *none_prefix[] = { NULL };
+static const char *creg_prefix[] = { "+CREG:", NULL };
+static const char *cops_prefix[] = { "+COPS:", NULL };
+static const char *csq_prefix[] = { "+CSQ:", NULL };
+static const char *cind_prefix[] = { "+CIND:", NULL };
+static const char *cmer_prefix[] = { "+CMER:", NULL };
+static const char *zpas_prefix[] = { "+ZPAS:", NULL };
+static const char *option_tech_prefix[] = { "_OCTI:", "_OUWCTI:", NULL };
+
+struct netreg_data {
[…]

### 2014-06-16 ofono-wifi/drivers/hfpmodem/hfpmodem.c

@@ -1,0 +1,58 @@
+/*
+ *
+ *  oFono - Open Source Telephony
+ *
+ *  Copyright (C) 2008-2011  Intel Corporation. All rights reserved.
+ *
+ *  This program is free software; you can redistribute it and/or modify
+ *  it under the terms of the GNU General Public License version 2 as
+ *  published by the Free Software Foundation.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU General Public License for more details.
+ *
+ *  You should have received a copy of the GNU General Public License
+ *  along with this program; if not, write to the Free Software
+ *  Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA  02110-1301  USA
+ *
+ */
+
+#ifdef HAVE_CONFIG_H
+#include <config.h>
+#endif
+
+#define OFONO_API_SUBJECT_TO_CHANGE
+#include <ofono/plugin.h>
+#include <ofono/modem.h>
+
+#include <glib.h>
+#include <gatchat.h>
+
+#include "hfpmodem.h"
+
+static int hfpmodem_init(void)
+{
+	hfp_voicecall_init();
+	hfp_devinfo_init();
+	hfp_netreg_init();
+	hfp_call_volume_init();
+	hfp_handsfree_init();
+	hfp_siri_init();
+
+	return 0;
+}
+
+static void hfpmodem_exit(void)
+{
+	hfp_voicecall_exit();
+	hfp_devinfo_exit();
+	hfp_netreg_exit();
+	hfp_call_volume_exit();
+	hfp_handsfree_exit();
+	hfp_siri_exit();
+}
+
+OFONO_PLUGIN_DEFINE(hfpmodem, "Hands-Free Profile Driver", VERSION,
+		OFONO_PLUGIN_PRIORITY_DEFAULT, hfpmodem_init, hfpmodem_exit)

### 2014-06-16 ofono-wifi/drivers/hfpmodem/slc.h

@@ -1,0 +1,48 @@
+/*
+ *
+ *  oFono - Open Source Telephony
+ *
+ *  Copyright (C) 2008-2011  Intel Corporation. All rights reserved.
+ *
+ *  This program is free software; you can redistribute it and/or modify
+ *  it under the terms of the GNU General Public License version 2 as
+ *  published by the Free Software Foundation.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU General Public License for more details.
+ *
+ *  You should have received a copy of the GNU General Public License
+ *  along with this program; if not, write to the Free Software
+ *  Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA  02110-1301  USA
+ *
+ */
+
+enum hfp_indicator {
+	HFP_INDICATOR_SERVICE = 0,
+	HFP_INDICATOR_CALL,
+	HFP_INDICATOR_CALLSETUP,
+	HFP_INDICATOR_CALLHELD,
+	HFP_INDICATOR_SIGNAL,
+	HFP_INDICATOR_ROAM,
+	HFP_INDICATOR_BATTCHG,
+	HFP_INDICATOR_LAST
+};
+
+typedef void (*hfp_slc_cb_t)(void *userdata);
+
+struct hfp_slc_info {
+	GAtChat *chat;
+	unsigned int ag_features;
+	unsigned int ag_mpty_features;
+	unsigned int hf_features;
+	unsigned char cind_pos[HFP_INDICATOR_LAST];
+	unsigned int cind_val[HFP_INDICATOR_LAST];
+};
+
+void hfp_slc_info_init(struct hfp_slc_info *info, guint16 version);
+void hfp_slc_info_free(struct hfp_slc_info *info);
+
+void hfp_slc_establish(struct hfp_slc_info *info, hfp_slc_cb_t connect_cb,
+				hfp_slc_cb_t failed_cb, void *userdata);

### 2014-06-16 ofono-wifi/gatchat/gatchat.c

@@ -1,0 +1,1573 @@
+/*
+ *
+ *  AT chat library with GLib integration
+ *
+ *  Copyright (C) 2008-2011  Intel Corporation. All rights reserved.
+ *
+ *  This program is free software; you can redistribute it and/or modify
+ *  it under the terms of the GNU General Public License version 2 as
+ *  published by the Free Software Foundation.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU General Public License for more details.
+ *
+ *  You should have received a copy of the GNU General Public License
+ *  along with this program; if not, write to the Free Software
+ *  Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA  02110-1301  USA
+ *
+ */
+
+#ifdef HAVE_CONFIG_H
+#include <config.h>
+#endif
+
+#include <stdio.h>
+#include <unistd.h>
+#include <string.h>
+#include <assert.h>
+#include <ctype.h>
+#include <errno.h>
+
+#include <glib.h>
+
+#include "ringbuffer.h"
+#include "gatchat.h"
+#include "gatio.h"
+
+/* #define WRITE_SCHEDULER_DEBUG 1 */
+
+#define COMMAND_FLAG_EXPECT_PDU			0x1
+#define COMMAND_FLAG_EXPECT_SHORT_PROMPT	0x2
+
+struct at_chat;
+static void chat_wakeup_writer(struct at_chat *chat);
+
+static const char *none_prefix[] = { NULL };
+
+struct at_command {
+	char *cmd;
+	char **prefixes;
+	guint flags;
+	guint id;
+	guint gid;
+	GAtResultFunc callback;
+	GAtNotifyFunc listing;
+	gpointer user_data;
+	GDestroyNotify notify;
+};
+
+struct at_notify_node {
+	guint id;
+	guint gid;
+	GAtNotifyFunc callback;
+	gpointer user_data;
+	GDestroyNotify notify;
+	gboolean destroyed;
+};
+
+typedef gboolean (*node_remove_func)(struct at_notify_node *node,
+					gpointer user_data);
+
+struct at_notify {
+	GSList *nodes;
+	gboolean pdu;
+};
+
+struct at_chat {
[…]

### 2014-06-16 ofono-wifi/plugins/mbpi.c

@@ -1,0 +1,610 @@
+/*
+ *
+ *  oFono - Open Source Telephony
+ *
+ *  Copyright (C) 2008-2011  Intel Corporation. All rights reserved.
+ *
+ *  This program is free software; you can redistribute it and/or modify
+ *  it under the terms of the GNU General Public License version 2 as
+ *  published by the Free Software Foundation.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU General Public License for more details.
+ *
+ *  You should have received a copy of the GNU General Public License
+ *  along with this program; if not, write to the Free Software
+ *  Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA  02110-1301  USA
+ *
+ */
+
+#ifdef HAVE_CONFIG_H
+#include <config.h>
+#endif
+
+#include <string.h>
+#include <fcntl.h>
+#include <sys/mman.h>
+#include <sys/stat.h>
+#include <sys/types.h>
+#include <errno.h>
+#include <unistd.h>
+
+#include <glib.h>
+
+#define OFONO_API_SUBJECT_TO_CHANGE
+#include <ofono/modem.h>
+#include <ofono/gprs-provision.h>
+
+#ifndef MBPI_DATABASE
+#define MBPI_DATABASE  "/usr/share/mobile-broadband-provider-info/" \
+							"serviceproviders.xml"
+#endif
+
+#include "mbpi.h"
+
+#define _(x) case x: return (#x)
+
+enum MBPI_ERROR {
+	MBPI_ERROR_DUPLICATE,
+};
+
+struct gsm_data {
+	const char *match_mcc;
+	const char *match_mnc;
+	GSList *apns;
+	gboolean match_found;
+	gboolean allow_duplicates;
+};
+
+struct cdma_data {
+	const char *match_sid;
+	char *provider_name;
+	gboolean match_found;
+};
+
+const char *mbpi_ap_type(enum ofono_gprs_context_type type)
+{
+	switch (type) {
+		_(OFONO_GPRS_CONTEXT_TYPE_ANY);
+		_(OFONO_GPRS_CONTEXT_TYPE_INTERNET);
[…]

### 2014-06-16 ofono-wifi/plugins/telit.c

@@ -1,0 +1,728 @@
+/*
+ *
+ *  oFono - Open Source Telephony
+ *
+ *  Copyright (C) 2008-2011  Intel Corporation. All rights reserved.
+ *
+ *  This program is free software; you can redistribute it and/or modify
+ *  it under the terms of the GNU General Public License version 2 as
+ *  published by the Free Software Foundation.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU General Public License for more details.
+ *
+ *  You should have received a copy of the GNU General Public License
+ *  along with this program; if not, write to the Free Software
+ *  Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA  02110-1301  USA
+ *
+ */
+
+#ifdef HAVE_CONFIG_H
+#include <config.h>
+#endif
+
+#include <errno.h>
+#include <stdlib.h>
+#include <stdio.h>
+#include <unistd.h>
+#include <fcntl.h>
+#include <termios.h>
+#include <string.h>
+#include <sys/socket.h>
+
+#include <glib.h>
+#include <gatchat.h>
+#include <gattty.h>
+
+#define OFONO_API_SUBJECT_TO_CHANGE
+#include <ofono/plugin.h>
+#include <ofono/log.h>
+#include <ofono/modem.h>
+#include <ofono/call-barring.h>
+#include <ofono/call-forwarding.h>
+#include <ofono/call-meter.h>
+#include <ofono/call-settings.h>
+#include <ofono/devinfo.h>
+#include <ofono/message-waiting.h>
+#include <ofono/netreg.h>
+#include <ofono/phonebook.h>
+#include <ofono/sim.h>
+#include <ofono/gprs.h>
+#include <ofono/gprs-context.h>
+#include <ofono/sms.h>
+#include <ofono/ussd.h>
+#include <ofono/voicecall.h>
+
+#include <drivers/atmodem/atutil.h>
+#include <drivers/atmodem/vendor.h>
+
+#include "bluez4.h"
+
+static const char *none_prefix[] = { NULL };
[…]

### 2014-06-16 ofono-wifi/src/dbus.c

@@ -1,0 +1,531 @@
+/*
+ *
+ *  oFono - Open Source Telephony
+ *
+ *  Copyright (C) 2008-2011  Intel Corporation. All rights reserved.
+ *
+ *  This program is free software; you can redistribute it and/or modify
+ *  it under the terms of the GNU General Public License version 2 as
+ *  published by the Free Software Foundation.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU General Public License for more details.
+ *
+ *  You should have received a copy of the GNU General Public License
+ *  along with this program; if not, write to the Free Software
+ *  Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA  02110-1301  USA
+ *
+ */
+
+#ifdef HAVE_CONFIG_H
+#include <config.h>
+#endif
+
+#include <glib.h>
+#include <gdbus.h>
+
+#include "ofono.h"
+
+#define OFONO_ERROR_INTERFACE "org.ofono.Error"
+
+static DBusConnection *g_connection;
+
+struct error_mapping_entry {
+	int error;
+	DBusMessage *(*ofono_error_func)(DBusMessage *);
+};
+
+struct error_mapping_entry cme_errors_mapping[] = {
+	{ 3,	__ofono_error_not_allowed },
+	{ 4,	__ofono_error_not_supported },
+	{ 16,	__ofono_error_incorrect_password },
+	{ 30,	__ofono_error_not_registered },
+	{ 31,	__ofono_error_timed_out },
+	{ 32,	__ofono_error_access_denied },
+	{ 50,	__ofono_error_invalid_args },
+};
+
+static void append_variant(DBusMessageIter *iter,
+				int type, const void *value)
+{
+	char sig[2];
+	DBusMessageIter valueiter;
+
+	sig[0] = type;
+	sig[1] = 0;
+
+	dbus_message_iter_open_container(iter, DBUS_TYPE_VARIANT,
+						sig, &valueiter);
+
+	dbus_message_iter_append_basic(&valueiter, type, value);
+
+	dbus_message_iter_close_container(iter, &valueiter);
+}
+
+void ofono_dbus_dict_append(DBusMessageIter *dict,
[…]

### 2014-06-16 ofono-wifi/src/handsfree.c

@@ -1,0 +1,629 @@
+/*
+ *
+ *  oFono - Open Source Telephony
+ *
+ *  Copyright (C) 2008-2011  Intel Corporation. All rights reserved.
+ *  Copyright (C) 2011  BMW Car IT GmbH. All rights reserved.
+ *
+ *  This program is free software; you can redistribute it and/or modify
+ *  it under the terms of the GNU General Public License version 2 as
+ *  published by the Free Software Foundation.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU General Public License for more details.
+ *
+ *  You should have received a copy of the GNU General Public License
+ *  along with this program; if not, write to the Free Software
+ *  Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA  02110-1301  USA
+ *
+ */
+
+#ifdef HAVE_CONFIG_H
+#include <config.h>
+#endif
+
+#define _GNU_SOURCE
+#include <string.h>
+#include <stdlib.h>
+#include <stdio.h>
+#include <errno.h>
+#include <unistd.h>
+
+#include <glib.h>
+
+#include <ofono/log.h>
+#include <ofono/modem.h>
+#include <ofono/handsfree.h>
+
+#include <gdbus.h>
+#include "ofono.h"
+#include "common.h"
+#include "hfp.h"
+
+static GSList *g_drivers = NULL;
+
+#define HANDSFREE_FLAG_CACHED 0x1
+
+struct ofono_handsfree {
+	ofono_bool_t nrec;
+	ofono_bool_t inband_ringing;
+	ofono_bool_t voice_recognition;
+	ofono_bool_t voice_recognition_pending;
+	unsigned int ag_features;
+	unsigned int ag_chld_features;
+	unsigned char battchg;
+	GSList *subscriber_numbers;
+
+	const struct ofono_handsfree_driver *driver;
+	void *driver_data;
+	struct ofono_atom *atom;
+	DBusMessage *pending;
+	int flags;
+};
+
+static const char **ag_features_list(unsigned int features,
[…]

### 2014-06-16 ofono-wifi/src/idmap.c

@@ -1,0 +1,202 @@
+/*
+ *
+ *  oFono - Open Source Telephony
+ *
+ *  Copyright (C) 2008-2011  Intel Corporation. All rights reserved.
+ *  Copyright (C) 2004  Red Hat, Inc. All Rights Reserved.
+ *
+ *  This program is free software; you can redistribute it and/or modify
+ *  it under the terms of the GNU General Public License version 2 as
+ *  published by the Free Software Foundation.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU General Public License for more details.
+ *
+ *  You should have received a copy of the GNU General Public License
+ *  along with this program; if not, write to the Free Software
+ *  Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA  02110-1301  USA
+ *
+ */
+
+#ifdef HAVE_CONFIG_H
+#include <config.h>
+#endif
+
+#define _GNU_SOURCE
+#include <string.h>
+
+#include <glib.h>
+
+#include "idmap.h"
+
+#define BITS_PER_LONG (sizeof(unsigned long) * 8)
+
+struct idmap {
+	unsigned long *bits;
+	unsigned int size;
+	unsigned int min;
+	unsigned int max;
+};
+
+static inline int ffz(unsigned long word)
+{
+	return __builtin_ctzl(~word);
+}
+
+/*
+ * Stolen from linux kernel lib/find_next_bit.c
+ */
+static unsigned int find_next_zero_bit(const unsigned long *addr,
+					unsigned int size,
+					unsigned int offset)
+{
+	const unsigned long *p = addr + offset / BITS_PER_LONG;
+	unsigned int result = offset & ~(BITS_PER_LONG-1);
+	unsigned long tmp;
+
+	if (offset >= size)
+		return size;
+
+	size -= result;
+	offset %= BITS_PER_LONG;
+
+	if (offset) {
+		tmp = *(p++);
+		tmp |= ~0UL >> (BITS_PER_LONG - offset);
+
+		if (size < BITS_PER_LONG)
+			goto found_first;
+
+		if (~tmp)
[…]

### 2014-06-16 ofono-wifi/unit/test-caif.c

@@ -1,0 +1,173 @@
+/*
+ *
+ *  oFono - Open Source Telephony
+ *
+ *  Copyright (C) 2008-2011  Intel Corporation. All rights reserved.
+ *
+ *  This program is free software; you can redistribute it and/or modify
+ *  it under the terms of the GNU General Public License version 2 as
+ *  published by the Free Software Foundation.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU General Public License for more details.
+ *
+ *  You should have received a copy of the GNU General Public License
+ *  along with this program; if not, write to the Free Software
+ *  Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA  02110-1301  USA
+ *
+ */
+
+#ifdef HAVE_CONFIG_H
+#include <config.h>
+#endif
+
+#include <errno.h>
+#include <fcntl.h>
+#include <unistd.h>
+#include <stdlib.h>
+#include <string.h>
+
+#pragma GCC diagnostic ignored "-Wdeprecated-declarations"
+
+#include <glib.h>
+#include <glib/gprintf.h>
+
+#include <gatchat.h>
+
+#include <drivers/stemodem/caif_socket.h>
+#include <drivers/stemodem/if_caif.h>
+
+static GMainLoop *mainloop;
+
+static int do_open(void)
+{
+	int fd;
+
+	fd = open("/dev/chnlat11", O_RDWR);
+	if (fd < 0) {
+		g_printerr("Open of chnlat11 failed (%d)\n", errno);
+		return -EIO;
+	}
+
+	return fd;
+}
+
+static int do_connect(void)
+{
+	struct sockaddr_caif addr;
+	int sk, err;
+
+	/* Create a CAIF socket for AT Service */
+	sk = socket(AF_CAIF, SOCK_SEQPACKET, CAIFPROTO_AT);
+	if (sk < 0) {
+		g_printerr("CAIF socket creation failed (%d)\n", errno);
+		return -EIO;
+	}
+
+	memset(&addr, 0, sizeof(addr));
+	addr.family = AF_CAIF;
[…]

### 2014-06-16 ofono-wifi/unit/test-mux.c

@@ -1,0 +1,570 @@
+/*
+ *
+ *  oFono - Open Source Telephony
+ *
+ *  Copyright (C) 2008-2011  Intel Corporation. All rights reserved.
+ *
+ *  This program is free software; you can redistribute it and/or modify
+ *  it under the terms of the GNU General Public License version 2 as
+ *  published by the Free Software Foundation.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU General Public License for more details.
+ *
+ *  You should have received a copy of the GNU General Public License
+ *  along with this program; if not, write to the Free Software
+ *  Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA  02110-1301  USA
+ *
+ */
+
+#ifdef HAVE_CONFIG_H
+#include <config.h>
+#endif
+
+#include <unistd.h>
+#include <stdlib.h>
+#include <string.h>
+#include <sys/socket.h>
+#include <netinet/in.h>
+#include <arpa/inet.h>
+
+#pragma GCC diagnostic ignored "-Wdeprecated-declarations"
+
+#include <glib.h>
+#include <glib/gprintf.h>
+
+#include "gatmux.h"
+#include "gsm0710.h"
+
+static int do_connect(const char *address, unsigned short port)
+{
+	struct sockaddr_in addr;
+	int sk, err;
+
+	sk = socket(PF_INET, SOCK_STREAM, 0);
+	if (sk < 0)
+		return sk;
+
+	memset(&addr, 0, sizeof(addr));
+	addr.sin_family = AF_INET;
+	addr.sin_addr.s_addr = inet_addr(address);
+	addr.sin_port = htons(port);
+
+	err = connect(sk, (struct sockaddr *) &addr, sizeof(addr));
[…]

### 2014-06-16 ofono-wifi/unit/test-simutil.c

@@ -1,0 +1,519 @@
+/*
+ *
+ *  oFono - Open Source Telephony
+ *
+ *  Copyright (C) 2008-2011  Intel Corporation. All rights reserved.
+ *
+ *  This program is free software; you can redistribute it and/or modify
+ *  it under the terms of the GNU General Public License version 2 as
+ *  published by the Free Software Foundation.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU General Public License for more details.
+ *
+ *  You should have received a copy of the GNU General Public License
+ *  along with this program; if not, write to the Free Software
+ *  Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA  02110-1301  USA
+ *
+ */
+
+#ifdef HAVE_CONFIG_H
+#include <config.h>
+#endif
+
+#include <string.h>
+#include <stdio.h>
+#include <assert.h>
+#include <glib.h>
+
+#include <ofono/types.h>
+
+#include "simutil.h"
+#include "util.h"
+
+/* Taken from 51.011 Appendix K.2 */
+const unsigned char valid_mms_params[] = {
+	0xAB, 0x81, 0x88, 0x80, 0x01, 0x01, 0x81, 0x17, 0x68, 0x74, 0x74, 0x70,
+	0x3A, 0x2F, 0x2F, 0x6D, 0x6D, 0x73, 0x2D, 0x6F, 0x70, 0x65, 0x72, 0x61,
+	0x74, 0x6F, 0x72, 0x2E, 0x63, 0x6F, 0x6D, 0x82, 0x32, 0x10, 0xAA, 0x08,
+	0x2B, 0x34, 0x39, 0x35, 0x33, 0x34, 0x31, 0x39, 0x30, 0x36, 0x00, 0x09,
+	0x87, 0x25, 0xC5, 0x0A, 0x90, 0x0C, 0x9A, 0x0D, 0x64, 0x75, 0x6D, 0x6D,
+	0x79, 0x5F, 0x6E, 0x61, 0x6D, 0x65, 0x00, 0x0E, 0x64, 0x75, 0x6D, 0x6D,
[…]

### 2014-06-16 ofono-wifi/plugins/telit.c
applied patch to make telit plugin talk over dbus to our relay daemon instead of bluetooth

@@ -33,6 +33,9 @@
 #include <sys/socket.h>
 
 #include <glib.h>
+#include <glib-object.h>
+#include <gdbus.h>
+// #include <gio/gio.h>
 #include <gatchat.h>
 #include <gattty.h>
 
@@ -63,6 +66,13 @@
 static const char *none_prefix[] = { NULL };
 static const char *rsen_prefix[]= { "#RSEN:", NULL };
 
+
+#define	SMARTE_SERVICE "org.smart_e.RSAP"
+#define	SMARTE_INTERFACE		SMARTE_SERVICE ".RSAPServer"
+// static GDBusProxy *dbProxy;
+// static GDBusClient *cardReader;
+static DBusConnection *connection;
+
 struct telit_data {
 	GAtChat *chat;		/* AT chat */
 	GAtChat *modem;		/* Data port */
@@ -84,13 +94,13 @@ static void telit_debug(const char *str, void *user_data)
 }
 
 static void hex_print(const char *buf, const gsize bytes_read) {
-    char buf_str[300 * 3 + 1] = { 0 };
-    char *endBuf = buf_str;
-    int i;
-    for (i = 0; i < bytes_read; i++) {
-        endBuf += sprintf(endBuf, "%02X ", (unsigned char)buf[i]);
-    }
-    DBG("read %zu bytes: %s", bytes_read, buf_str);
+	char buf_str[300 * 3 + 1] = { 0 };
+	char *endBuf = buf_str;
+	int i;
+	for (i = 0; i < bytes_read; i++) {
+		endBuf += sprintf(endBuf, "%02X ", (unsigned char)buf[i]);
+	}
+	DBG("read %zu bytes: %s", bytes_read, buf_str);
 }
 
 static void sap_close_io(struct ofono_modem *modem)
@@ -115,51 +125,6 @@ static void sap_close_io(struct ofono_modem *modem)
 		g_source_remove(data->hw_watch);
 }
 
-static void bt_watch_remove(gpointer userdata)
-{
-	struct ofono_modem *modem = userdata;
-	struct telit_data *data = ofono_modem_get_data(modem);
-
-	ofono_modem_set_powered(modem, FALSE);
-
-	data->bt_watch = 0;
-}
-
-static gboolean bt_event_cb(GIOChannel *bt_io, GIOCondition condition,
-							gpointer userdata)
-{
-	struct ofono_modem *modem = userdata;
[…]

### 2014-06-16 ofono-wifi/plugins/telit.c
some debug output

@@ -175,9 +175,12 @@ static gboolean smartcard_cb (DBusPendingCall *call, gpointer userdata)
     
     // hex_print(smartcard_response, bytes_read);
     GIOStatus status;
-    if (bytes_read > 0) 
+    if (bytes_read > 0) {
 		status = g_io_channel_write_chars(data->hw_io, smartcard_response,
 					bytes_read, &bytes_written, NULL);
+		DBG("smartcard event <--- (read %d, wrote %d bytes to hw_io)", (int)bytes_read, (int)bytes_written);
+		hex_print(buf, bytes_read);
+    }
 	if (status != G_IO_STATUS_NORMAL && status != G_IO_STATUS_AGAIN)
 		return FALSE;
 	// DBG("wrote %zu bytes back to hw_io", bytes_written);
@@ -202,12 +205,8 @@ static gboolean hw_event_cb(GIOChannel *hw_io, GIOCondition condition,
 							&bytes_read, NULL);
 
 		if (bytes_read > 0) {
-      		// hex_print(buf, bytes_read);
-    	}
-
-
-		if (bytes_read > 0) {
-			// DBG("some bytes read, rewriting");
+			DBG("some bytes read, rewriting");
+			hex_print(buf, bytes_read);
 			DBusPendingCall *call;
 			DBusMessage* processAPDUCall = dbus_message_new_method_call("org.smart_e.RSAP", "/RSAPServer", 
 				"org.smart_e.RSAPServer", "processAPDU");
@@ -522,7 +521,7 @@ static void rsen_disable_cb(gboolean ok, GAtResult *result, gpointer user_data)
 
 static int telit_sap_open(void)
 {
-	const char *device = "/dev/ttyUSB4";
+	const char *device = "/dev/ttyUSB0";
 	DBG("");
 	struct termios ti;
 	int fd;

### 2014-06-16 rpi-sim/dbusrelaynet.py

@@ -74,7 +74,7 @@ class Server(dbus.service.Object):
     @dbus.service.method("org.smart_e.RSAPServer",
                          in_signature='ay', out_signature='ay')
     def processAPDU(self, inCommand):
-        print 'HARDWARE > ', toHexString(list(bytearray(inCommand)))
+        print 'INCOMING > ', toHexString(list(bytearray(inCommand)))
         relay = []
         for i in inCommand:
             num = i & 0xFF
@@ -96,11 +96,11 @@ class Server(dbus.service.Object):
                 self.respQueue.put(resp)
 
     def process(self, inCommand):
-        print "<", toHexString(inCommand)
+        # print "<", toHexString(inCommand)
         self.client.send_json(inCommand)
         apduResponse = self.client.recv_json()
         bArrResp = list(bytearray(apduResponse))
-        print ">", toHexString(bArrResp)
+        print "OUTGOING > ", toHexString(bArrResp)
         return apduResponse
 
 

### 2014-06-16 ofono-wifi/plugins/sap.c
also applied the sap plugin changes

@@ -120,74 +120,37 @@ static void sap_remove(struct ofono_modem *modem)
 	ofono_modem_set_data(modem, NULL);
 }
 
-static void sap_connect_reply(DBusPendingCall *call, gpointer user_data)
-{
-	struct ofono_modem *modem = user_data;
-	struct sap_data *data = ofono_modem_get_data(modem);
-	DBusError derr;
-	DBusMessage *reply;
-	int fd, err;
-
-	DBG("");
-
-	reply = dbus_pending_call_steal_reply(call);
-
-	data->call = NULL;
-
-	if (ofono_modem_get_powered(modem))
-		goto done;
-
-	dbus_error_init(&derr);
-	if (dbus_set_error_from_message(&derr, reply)) {
-
-		DBG("Connect reply: %s", derr.message);
-
-		dbus_error_free(&derr);
-		goto done;
-	}
-
-	if (!dbus_message_get_args(reply, NULL, DBUS_TYPE_UNIX_FD, &fd,
-				DBUS_TYPE_INVALID))
-		goto done;
-
-	data->hw_modem = sap_hw_modem;
-	data->sap_driver = sap_hw_driver;
-
-	err = data->sap_driver->enable(data->hw_modem, modem, fd);
-	if (!err || err == -EINPROGRESS) {
-		dbus_message_unref(reply);
-		return;
-	}
-
-done:
-	ofono_modem_set_powered(modem, FALSE);
-	dbus_message_unref(reply);
-}
-
-/* power up hardware */
 static int sap_enable(struct ofono_modem *modem)
 {
 	struct sap_data *data = ofono_modem_get_data(modem);
 	DBusPendingCall *call;
 	int status;
+	int fd, err;
 	const char *str = "sap";
 	const char *server_path = ofono_modem_get_string(modem, "ServerPath");
 
-	DBG("%p", modem);
+	DBG("%p (hw %p)", modem, sap_hw_modem);
+	DBG("non-Bluetooth version");
 
-	status = bluetooth_send_with_reply(server_path, BLUEZ_SERIAL_INTERFACE,
-					"ConnectFD", &call, sap_connect_reply,
-					modem, NULL, DBUS_TIMEOUT,
-					DBUS_TYPE_STRING, &str,
-					DBUS_TYPE_INVALID);
+	if (ofono_modem_get_powered(modem))
+		goto done;
 
-	if (status < 0)
-		return -EINVAL;
+	
+	data->hw_modem = sap_hw_modem;
[…]

### 2014-06-16 ofono-wifi/plugins/telit.c
working version over wifi 16/02/14

@@ -179,7 +179,7 @@ static gboolean smartcard_cb (DBusPendingCall *call, gpointer userdata)
 		status = g_io_channel_write_chars(data->hw_io, smartcard_response,
 					bytes_read, &bytes_written, NULL);
 		DBG("smartcard event <--- (read %d, wrote %d bytes to hw_io)", (int)bytes_read, (int)bytes_written);
-		hex_print(buf, bytes_read);
+		hex_print(smartcard_response, bytes_read);
     }
 	if (status != G_IO_STATUS_NORMAL && status != G_IO_STATUS_AGAIN)
 		return FALSE;

### 2014-06-16 ofono-wifi/src/modem.c

@@ -2005,7 +2005,7 @@ int ofono_modem_register(struct ofono_modem *modem)
 	modem->online_watches = __ofono_watchlist_new(g_free);
 	modem->powered_watches = __ofono_watchlist_new(g_free);
 
-	emit_modem_added(modem);
+	// emit_modem_added(modem);
 	call_modemwatches(modem, TRUE);
 
 	modem->sim_watch = __ofono_modem_add_atom_watch(modem,

### 2014-06-16 ofono-bt/plugins/telit.c
not writing CONNECT_RESP from bluetooth to hw_io makes modem work correctly over by

@@ -141,6 +141,10 @@ static gboolean bt_event_cb(GIOChannel *bt_io, GIOCondition condition,
 
 		if (bytes_read > 0) {
 			GIOStatus st;
+			if (bytes_read >= 3 && buf[0] == 0x01 && buf[1] == 0x01 && buf[2] == 0x00) {
+				DBG("do not write CONNECT_RESP");
+				return TRUE;	
+			}
 			st = g_io_channel_write_chars(data->hw_io, buf,
 					bytes_read, &bytes_written, NULL);
 			DBG("BT event <--- (read %d, wrote %d bytes to hw_io)", (int)bytes_read, (int)bytes_written);

### 2014-06-16 ofono-bt/src/modem.c

@@ -2005,7 +2005,7 @@ int ofono_modem_register(struct ofono_modem *modem)
 	modem->online_watches = __ofono_watchlist_new(g_free);
 	modem->powered_watches = __ofono_watchlist_new(g_free);
 
-	emit_modem_added(modem);
+	//emit_modem_added(modem);
 	call_modemwatches(modem, TRUE);
 
 	modem->sim_watch = __ofono_modem_add_atom_watch(modem,

### 2014-06-17 paper.tex
a round of edditing, fitting to 3 pages

@@
 
-Secondly, the price is largely driven by the high infrastructure costs which increase with the growing number of network users - the increasing number of connected clients periodically communicating data also increase cellular networks loads, in the extreme cases causing \emph{signaling storms}~\cite{Yang:wh} that can bring the entire network down.
+Secondly, the price is largely driven by the high infrastructure costs which increase with the growing number of network users - the increasing number of connected, periodically communicating clients also increases cellular network load, in the extreme cases causing \emph{signaling storms}~\cite{Yang:wh} that can bring down an entire network.
 
 Finally, energy consumption of always connected clients is high compared to short-range radio technologies.
-We have analyzed in the past how power-hungry such communication can be from the mobile device perspective~\cite{Aucinas:2013uk} and without careful optimization of such communication patterns the cost may be too high for many types of devices.
+We have analyzed in the past power consumption of cellular communication~\cite{Aucinas:2013uk} and careful communication optimization is necessary for battery-powered devices.
 
-These factors all motivate the need for \emph{PhoneLets}, a wireless system that enables offloading of cellular connectivity off a device to share a subscription between devices, balance energy consumption of a group of devices or simply replace the cellular link with a short-range link to a powered device.
+These factors all motivate the need for \emph{PhoneLets}, a wireless system that enables offloading of cellular connectivity off a device to share subscriber identity between multiple devices, balance energy consumption of a group of devices or replace the cellular link with a short-range link to a powered device.
 
 \emph{PhoneLets} enable the functionality by exploiting remote SIM functionality of certain hardware, whereby a cellular module accepts to exchange commands with a SIM/UICC module over a separate communication link rather than direct, physical connection.
-Building our prototype on the existing technology allows us to achieve more advanced functionality without breaking encryption of participating entities or doing unlawful communication interception.
-Instead, \emph{PhoneLets} infrastructure establishes and manages secure communication channels between participating entities and allows a device to connect to a cellular network without a physical SIM module being present.
+\emph{PhoneLets} infrastructure establishes and manages secure communication channels between participating entities and allows a device to connect to a cellular network without a physical SIM module being present.
 
@@
 \centering
-\includegraphics[width=0.95\columnwidth]{figs/arch}
+\includegraphics[width=0.9\columnwidth]{figs/arch}
+\vspace{-3mm}
 \caption{Architecture}
 \label{fig:arch}
+\vspace{-5mm}
 \end{figure}
@@
 The overall concept and system architecture is shown in Fig.~\ref{fig:arch}.
-There are three types of entities: mobile server - a mobile device which can share its SIM with other clients, SIM server which may host and provide more sets of credentials over local or wide-area networks and a mobile client which can use either of the methods.
+There are three types of entities: 1) mobile server, a mobile device which can share its SIM with other clients, 2) SIM server which may host and provide more sets of credentials over local or wide-area networks and 3) a mobile client which can use either of the methods.
 
-Atypically for cellular networking technologies, certain parts are relatively loosely coupled as the baseband does not require physical presence of a SIM  or relays communication to its SIM.
-Furthermore the SIM is a smartcard which can be interfaced with a standard smartcard reader, allowing for a relatively simple server design.
+Some of the components are only loosely coupled as our selected module supports the standardized remote SIM protocol version.
+The SIM on the other hand is a smartcard which can be interfaced with a standard smartcard reader, allowing for a relatively simple server design.
 Our work for this demo focused primarily on the client and SIM server design, and we are using off-the-shelf hardware as the mobile server to show compatibility with the specification.
@@
 \centering
-\includegraphics[width=0.95\columnwidth]{figs/client}
+\includegraphics[width=0.9\columnwidth]{figs/client}
+\vspace{-2mm}
 \caption{Client prototype}
 \label{fig:client}
+\vspace{-3mm}
 \end{figure}
@@
 
-In addition to an unmodified Galaxy S4 running as the mobile client, we have also implemented a server that does not require a cellular module and instead interfaces with SIM cards through a smartcard reader.
-Once a connection is established to either backend (over BlueTooth to the mobile server or an encrypted connection to the SIM server), the cellular module communicates with the remote card using the standard Application Protocol Data Unit (APDU) messages (Fig.~\ref{fig:sequence}) and registers with the network.
-Importantly, in this case we replace the BlueTooth communication channel with an arguably more secure option in that we use certificate-based authentication of both client and server and use strong SSL/TLS encryption for all data exchange.
+\begin{figure}[t!]
+\centering
+\includegraphics[width=0.9\columnwidth]{figs/sequence}
+\vspace{-2mm}
+\caption{Communication sequence between client, server and remote SIM card}
+\label{fig:sequence}
+\vspace{-5mm}
+\end{figure}
 
-An important point to make is that we do not clone the SIM card and by design the system disables connectivity for one device before enabling it for another one to prevent multiple devices sharing the same credentials simultaneously as this would very likely be unsupported by the network even if it were not considered to be an attempt to illegally clone a card.
-Therefore, one device is connected to the cellular network and may share (\emph{tether}) the connectivity to other devices over another channel.
-One cellular module is also only capable of being registered with one network, therefore enforcing a strict one-to-one mapping.
+In addition to an unmodified Galaxy S4 running as the mobile server, we have also implemented a server that only interfaces with SIM cards through a smartcard reader.
+Once a connection is established to either backend (over BlueTooth to the mobile server or an encrypted network connection to the SIM server), the cellular module communicates with the remote card using standard Application Protocol Data Unit (APDU) messages (Fig.~\ref{fig:sequence}) and registers with the network.
+In our design we replace the BlueTooth communication channel with an arguably more secure option in that we use certificate-based authentication of both client and server, and strong encryption for all data exchange.
+
+An important point is that we do not clone the SIM card and by design the system disables connectivity for one device before enabling it for another one to prevent multiple devices sharing the same credentials simultaneously.
+Therefore, only one device is connected to the cellular network and may share (\emph{tether}) the connectivity to other devices over another channel.
+One cellular module is also only capable of being registered with one network, enforcing a strict one-to-one mapping.
 
 One part of the network's security is that the network authenticates the subscriber for any major operation (e.g.
-placing or receiving phonecalls, establishing PDP context) as well as periodically checks for subscriber's presence, therefore the link between the client and the corresponding server must be always maintained, simplifying security requirements of the communication channel.
-
-\begin{figure}[t!]
-\centering
-\includegraphics[width=0.95\columnwidth]{figs/sequence}
-\caption{Communication sequence between client, server and remote SIM card}
-\label{fig:sequence}
-\end{figure}
+placing or receiving phonecalls, establishing PDP context), therefore the link between the client and the corresponding server must be always maintained, simplifying security requirements.
 
@@
   Apps background - WiFi    & 27    \\ \hline
-  BT SIM access             & 28    \\ \hline
-  WiFi frequent messages    & YY    \\ \hline
+  BlueTooth SIM access      & 28    \\ \hline
   Apps background - 3G      & 40    \\ \hline
@@
 }
+\vspace{-3mm}
 \caption{Energy consumption of the different usage scenarios of a mobile device.}
 \label{tab:energy}
+\vspace{-3mm}
 \end{table}
 
-We measured energy consumption of a Samsung Galaxy S4 phone using the Power Monitor\footnote{Monsoon Power Monitor \url{http://www.msoon.com/LabEquipment/PowerMonitor/}}.
+Energy consumption of a Samsung Galaxy S4 was measured\footnote{Using Monsoon Power Monitor \url{http://www.msoon.com/LabEquipment/PowerMonitor/}} in a number of scenarios.
 The different values are presented in Table.~\ref{tab:energy}.
@@
 
-With the default implementation of Remote SIM Access over BlueTooth, the device consumes three times the amount of energy of an idle device connected to a 3G network, which is a high energy cost to pay, however the amount is similar or even less than that of a device with multiple applications running in background but otherwise idle.
-Furthermore, energy consumption of WiFi tethering even when a client connected to it is idle is significantly higher than that of sharing connectivity by offloading the phone functionality.
+With the default implementation of Remote SIM Access over BlueTooth, the device consumes three times the amount of energy of an idle device connected to a 3G network, which is a high energy cost to pay, however the amount is comparable to that of an idle device with multiple applications running in background.
+
+Since we have implemented SIM sharing over WiFi using our custom client, we verified that the BlueTooth version is unnecessarily wasteful - once the PhoneLet has registered to the network all communication with the SIM server (for authentication) happens on demand, \eg when a new phonecall is received after 30 minutes of no activity.
+We therefore expect energy consumption of an optimized implementation over WiFi to be very close to that of an idle WiFi channel.
+For comparison, energy consumption of WiFi tethering even with a single idle client is significantly higher than that of sharing connectivity by offloading the phone functionality.
 
 These are early results, and we have not yet been able to replace the BlueTooth channel for an Android device.
-Nevertheless, the results point to a potential use-case where a powered \emph{PhoneLet} would be registered to the network on behalf of a WiFi-connected client with a considerable energy benefit: client connected over WiFi and exchanging messages of the same size and frequency as required by the network consumes YYmA of energy on average.
+Nevertheless, the results point to a potential use-case where a powered \emph{PhoneLet} would be registered to the network on behalf of a WiFi-connected client with a considerable energy benefit.
 The benefit is further improved when network access is shared across multiple such devices and SIM access load is balanced across them.
+The approach is highly suitable for providing wide-area network connectivity for a larger number of devices where other options are not available.f
 
@@
 
-To demonstrate the working prototype of the system, we show our client \emph{PhoneLet} (Fig.~\ref{fig:client}), communicating either with an off-the shelf mobile device, or a SIM server running on a raspberryPI to share their
-credentials to connect to their specific networks. We do so in three steps:
+To demonstrate the working prototype of the system, we show our client \emph{PhoneLet} (Fig.~\ref{fig:client}), communicating with an off-the shelf mobile device or our SIM server to share their credentials for network registration in the following steps:
 
@@
 
-In the three number of steps we show our prototype's ability to offload phone functionality off a phone - share subscriber credentials between multiple clients and select which one will be connected to the network.
+In the three steps we show our prototype's ability to offload phone functionality off a phone - share subscriber credentials between multiple clients and select which one will be connected to the network.
 
@@
 
-As much as the telephone number is important, multiple providers have offered a service in the past whereby multiple SIM cards would be associated with the number and multiple handsets would ring on an incoming call\footnote{Google have also offered such service with Google Voice \url{http://www.google.co.uk/googlevoice/about.html\#tab=oneNumber}}, however it does not solve the problem of an increasing number of cellular clients and billing depends on the exact agreement with the service provider.
-A cellular network in UK has even offered a service to make phonecalls from tablets and laptops over WiFi to its users using a special app\footnote{O2 TU Go \url{http://www.o2.co.uk/apps/tu-go}}, decoupling number from the phone even further.
+As much as the telephone number is important, multiple providers have offered a service in the past whereby multiple SIM cards would be associated with the number and multiple handsets would ring on an incoming call, \eg Google Voice, however it does not solve the problem of an increasing number of cellular clients and billing depends on the exact agreement with the service provider.
+One cellular network in UK has even offered a service to make phonecalls from tablets and laptops over WiFi to its users using a special app\footnote{O2 TU Go \url{http://www.o2.co.uk/apps/tu-go}}.
 
@@
 (2) \emph{Two standard UK power sockets}.
-(3) \emph{Two cellular network subscriptions (SIM cards)}, but are to be sources by the presenters, (4) an HDMI display during setup.
+(3) \emph{Two cellular network subscriptions (SIM cards)}, but are to be sources by the presenters, (4) WiFi AP for local communication.
 Setup time - up to an hour for connecting the demo up and verifying operation.

### 2014-06-17 paper.tex
changed a bit of discussion

@@
 
-\emph{PhoneLets} enable the functionality by exploiting remote SIM functionality of certain hardware, whereby a cellular module accepts to exchange commands with a SIM/UICC module over a separate communication link rather than direct, physical connection.
+\emph{PhoneLets} enable the functionality by exploiting remote SIM functionality of existing hardware, whereby a cellular module accepts to exchange commands with a SIM/UICC module over a separate communication link rather than direct, physical connection.
+Some uses of such technology in the past have included automotive as well as electronic currency~\cite{subramanian2003sim}.
 \emph{PhoneLets} infrastructure establishes and manages secure communication channels between participating entities and allows a device to connect to a cellular network without a physical SIM module being present.
@@
 \label{fig:arch}
-\vspace{-5mm}
+\vspace{-2mm}
 \end{figure}
@@
 Our demo mobile client (Fig.~\ref{fig:client}) consists of a Telit UC864 cellular module connected to a raspberryPI as a controller, which also uses WiFi and BlueTooth dongles for local wireless connectivity.
-The controller runs a modified version of the Ofono open telephony stack\footnote{\url{https://ofono.org/}} and manages communication between the cellular module and multiple backends.
+The controller runs a modified version of the Ofono open telephony stack and manages communication between the cellular module and multiple backends.
 
@@
 \label{fig:sequence}
-\vspace{-5mm}
+\vspace{-4mm}
 \end{figure}
@@
 \hline
-  \textbf{Use case}         & \textbf{Avg. energy consumption (mA)}  \\ \hline
+  \textbf{Use case}         & \textbf{Energy consumption (mA)}  \\ \hline
   Idle - airplane mode      & 5     \\ \hline
@@
 }
-\vspace{-3mm}
+\vspace{-2mm}
 \caption{Energy consumption of the different usage scenarios of a mobile device.}
@@
 Energy consumption of a Samsung Galaxy S4 was measured\footnote{Using Monsoon Power Monitor \url{http://www.msoon.com/LabEquipment/PowerMonitor/}} in a number of scenarios.
-The different values are presented in Table.~\ref{tab:energy}.
+The different values are presented in Table.~\ref{tab:energy}\footnote{Energy consumption provided as average current drawn because it is independent of current battery voltage}.
 Energy consumption of an idle device with applications in background (\emph{Apps background}) considers a device with Google Apps, and Facebook installed and running in background and is consistent with our earlier observations~\cite{Aucinas:2013uk}.
@@
     \item Start sharing its SIM card with the client and show that the client is registered to the network with the same ID by making a phone call to it
-    \item Start using SIM card in the SIM server and show that the client is registered to the network with a different ID by making a call to the other number.
+    \item Start using SIM card in the SIM server over WiFi and show that the client is registered to the network with a different ID by making a call to the other number.
 \end{enumerate}
@@
 
-As much as the telephone number is important, multiple providers have offered a service in the past whereby multiple SIM cards would be associated with the number and multiple handsets would ring on an incoming call, \eg Google Voice, however it does not solve the problem of an increasing number of cellular clients and billing depends on the exact agreement with the service provider.
+\emph{PhoneLets} also offers management of part of an individual's ID - the phone number.
+In the past, multiple providers have offered a service in the past whereby multiple SIM cards would be associated with the number and multiple handsets would ring on an incoming call, \eg Google Voice, however it does not solve the problem of an increasing number of cellular clients and billing depends on the exact agreement with the service provider.
 One cellular network in UK has even offered a service to make phonecalls from tablets and laptops over WiFi to its users using a special app\footnote{O2 TU Go \url{http://www.o2.co.uk/apps/tu-go}}.
@@
 
-Finally, security is a major concern and the reason why secure elements are used by operators in the first place despite multiple proposals to move towards software-based SIMs.
-In our proposal, however, we still maintain the separate secure element and only replace the interface to it with an arguably secure alternative widely used in industry.
-Furthermore, past proposals of remote SIM usage have already included mobile payment solutions~\cite{subramanian2003sim}, which are even more risky than communication alone.
-Detailed security analysis is subject to future work, but we believe that any additional risk created by \emph{PhoneLets} is marginal.
+Finally, questions that remain to be answered in future work include not only optimizations of the management and control protocols, but also the development of usage policies in multi-user environments and guaranteed isolation properties as well as adaptation of energy-aware routing protocols for distributed nodes.
 
@@
 
-(1) \emph{Desk/table of approximately 1.5m, standard-width} to accommodate the demo.
+(1) \emph{Desk/table of approximately 2m, standard-width} to accommodate the demo.
 (2) \emph{Two standard UK power sockets}.

### 2014-06-26 paper.tex
edits based on Toby's feedback

@@
 \begin{abstract}
-This demo presents \emph{PhoneLets}, a system to provide offloading of phone functionality of mobile cellular devices to a neighboring, potentially powered one.
-Phone offloading happens through providing access to cellular network subscriber credentials over a short-range wireless link and selecting the device that uses them to register to a particular network.
-As a result, we can use a single identity (SIM card) across multiple clients without increasing their connectivity costs, the load for the network handling large numbers of other subscribers and we are able to optimize mobile clients' energy consumption by leveraging powered \emph{PhoneLets} or using them to share the energy consumption load across distributed nodes.
+This demo presents how phone functionality can be offloaded from a smartphone over wireless link to a \emph{PhoneLet} by sharing one SIM card across multiple devices.
+This can lead to significant cost and network load reductions by reducing the number of simultaneously connected mobile clients.
+Furthermore, it can bring significant energy reductions for the mobile user when connected to a powered \emph{PhoneLet}.
+It absorbs the energy cost of inefficient mobile applications' communication patterns, instead providing connectivity for the user over a WiFi link.
 \end{abstract}
@@
 
-Not only the modern smartphone has more in common with a computer than a phone, but connectivity to the Internet as well as the increasing number of surrounding devices (the \emph{Internet of Things}) is tipping the balance even further.
-Cellular networks, however, are just as important as they provide almost ubiquitous Internet connectivity around the world.
-There are multiple problems with cellular connectivity.
+A modern smartphone has more in common with a computer than a phone, and connectivity to the Internet and surrounding devices (the \emph{Internet of Things}) is tipping the balance even further.
+Importance of cellular networks, however, remains as they provide almost ubiquitous Internet connectivity around the world.
+Cellular connectivity has its own problems.
+This demo shows how some of them can be mitigated by offloading cellular connectivity to a \emph{PhoneLet} - sharing one SIM (Subscriber Identity Module) between nearby devices.
 
 Firstly, it is relatively expensive for the users~\cite{Anonymous:2013ut} and typically the end user needs to obtain a separate subscription for each of his devices connected to the cellular network, e.g.
-smartphone and tabled as well as broadband connectivity at home and at work, each with its underutilized caps and subscription fees.
+smartphone and tablet as well as broadband connectivity at home and at work.
+Each device often has its own subscription and fees and sharing a single SIM across them solves the issue.
 
 Secondly, the price is largely driven by the high infrastructure costs which increase with the growing number of network users - the increasing number of connected, periodically communicating clients also increases cellular network load, in the extreme cases causing \emph{signaling storms}~\cite{Yang:wh} that can bring down an entire network.
+Our approach reduces the number of simultaneously connected clients by sharing the same cellular link across all clients.
 
 Finally, energy consumption of always connected clients is high compared to short-range radio technologies.
-We have analyzed in the past power consumption of cellular communication~\cite{Aucinas:2013uk} and careful communication optimization is necessary for battery-powered devices.
+Sharing a cellular link also helps reduce it for mobile devices.
+In previous work we analyzed the power consumption of cellular communication~\cite{Aucinas:2013uk} and showed that careful optimization is necessary to prolong the battery life of mobile devices.
+We discuss in Section~\ref{sec:energy} how PhoneLets can help with such optimization.
 
-These factors all motivate the need for \emph{PhoneLets}, a wireless system that enables offloading of cellular connectivity off a device to share subscriber identity between multiple devices, balance energy consumption of a group of devices or replace the cellular link with a short-range link to a powered device.
-
-\emph{PhoneLets} enable the functionality by exploiting remote SIM functionality of existing hardware, whereby a cellular module accepts to exchange commands with a SIM/UICC module over a separate communication link rather than direct, physical connection.
+\emph{PhoneLets} enable phone offloading by exploiting remote SIM (RSAP) functionality of existing hardware.
+A cellular module exchange commands with a SIM/UICC module over a wireless communication link rather than direct, physical connection.
 Some uses of such technology in the past have included automotive as well as electronic currency~\cite{subramanian2003sim}.
@@
 
-\section{Strawman Design}
+\section{PhoneLets Design}
 
@@
 
-Some of the components are only loosely coupled as our selected module supports the standardized remote SIM protocol version.
-The SIM on the other hand is a smartcard which can be interfaced with a standard smartcard reader, allowing for a relatively simple server design.
-Our work for this demo focused primarily on the client and SIM server design, and we are using off-the-shelf hardware as the mobile server to show compatibility with the specification.
+Our design extends the standardized RSAP version by using a SIM server which is not capable of mobile connectivity itself and by using arbitrary communication links to access a remote SIM card (\emph{no SIM} in the mobile client).
+Our work for this demo focused primarily on the client and SIM server design, and we are using off-the-shelf hardware as the mobile server to demonstrate backwards-compatibility.
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=0.9\columnwidth]{figs/sequence}
+\vspace{-2mm}
+\caption{Communication sequence between client, server and remote SIM card}
+\label{fig:sequence}
+\vspace{-4mm}
+\end{figure}
+
+Mobile Client can obtain network credentials from any server -- a user's mobile phone, a co-located or a remote SIM server -- as long as it can reach it.
+The key observation is that there are few operations that require access to the SIM card due to its interface speed (it would be infeasible to use it on the data path) and instead it is used to generate encryption keys.
+The generating keys, however, can never be accessed directly.
+Part of the network's security however comes from the network authenticating the subscriber for any major operation (e.g.
+placing or receiving calls, establishing PDP context).
+The cellular module therefore must always be able to access the remote card using standard Application Protocol Data Unit (APDU) messages (Fig.~\ref{fig:sequence}).
+
+Importantly, we do not clone the SIM card and by design the system disables connectivity for one device before enabling it for another one to prevent multiple devices sharing the same credentials simultaneously.
+Only one device is connected to the cellular network and may share (\emph{tether}) the connectivity to other devices over another channel.
+One cellular module is also only capable of being registered with one network, enforcing a strict one-to-one mapping.
+
+\subsection{Prototype}
 
@@
 
-Our demo mobile client (Fig.~\ref{fig:client}) consists of a Telit UC864 cellular module connected to a raspberryPI as a controller, which also uses WiFi and BlueTooth dongles for local wireless connectivity.
+Our demo mobile client (Fig.~\ref{fig:client}), the \emph{PhoneLet}, consists of a Telit UC864 cellular module connected to a raspberryPI as a controller, which also uses WiFi and BlueTooth dongles for local wireless connectivity.
 The controller runs a modified version of the Ofono open telephony stack and manages communication between the cellular module and multiple backends.
 
-\begin{figure}[t!]
-\centering
-\includegraphics[width=0.9\columnwidth]{figs/sequence}
-\vspace{-2mm}
-\caption{Communication sequence between client, server and remote SIM card}
-\label{fig:sequence}
-\vspace{-4mm}
-\end{figure}
-
-In addition to an unmodified Galaxy S4 running as the mobile server, we have also implemented a server that only interfaces with SIM cards through a smartcard reader.
-Once a connection is established to either backend (over BlueTooth to the mobile server or an encrypted network connection to the SIM server), the cellular module communicates with the remote card using standard Application Protocol Data Unit (APDU) messages (Fig.~\ref{fig:sequence}) and registers with the network.
-In our design we replace the BlueTooth communication channel with an arguably more secure option in that we use certificate-based authentication of both client and server, and strong encryption for all data exchange.
-
-An important point is that we do not clone the SIM card and by design the system disables connectivity for one device before enabling it for another one to prevent multiple devices sharing the same credentials simultaneously.
-Therefore, only one device is connected to the cellular network and may share (\emph{tether}) the connectivity to other devices over another channel.
-One cellular module is also only capable of being registered with one network, enforcing a strict one-to-one mapping.
-
-One part of the network's security is that the network authenticates the subscriber for any major operation (e.g.
-placing or receiving phonecalls, establishing PDP context), therefore the link between the client and the corresponding server must be always maintained, simplifying security requirements.
+The two servers are an off-the-shelf Samsung Galaxy S4 and our custom server that only interfaces with SIM cards through a smartcard reader, also running on a raspberryPI.
+In our prototype, the server is connected to a WiFi AP to be reachable by the mobile client.
+We replace the BlueTooth communication channel with a more secure option - TCP link with certificate-based authentication of both client and server, and strong encryption for all data exchange.
+Once the client establishes a connection to either backend, it registers with the network and takes on the role of a phone.
 
 \section{Energy saving opportunities}
+\label{sec:energy}
 
-A major motivation for using \emph{PhoneLets} is reduced mobile client energy consumption.
-In this section we provide energy consumption of a few example scenarios using our early prototype.
-Since the client itself is not yet energy-optimized, i.e.
-does not use the full energy saving capabilities of the radio module and the controlling RaspberryPI is not meant for mobile uses, we focused on energy comparison of the mobile SIM server.
+Cost and network resource savings of using one active subscriber instead of multiple are straightforward, but reduced mobile client energy consumption is an important motivation for \emph{PhoneLets}.
+We analyze energy consumption of the client in a few example scenarios using our early prototype.
+The \emph{PhoneLet} itself is not yet energy-optimized, i.e.
+the controlling RaspberryPI is not meant for mobile uses, we focused on energy comparison of the mobile SIM server - a phone carried by the user from which we offload phone functionality.
 
@@
 \vspace{-2mm}
-\caption{Energy consumption of the different usage scenarios of a mobile device.}
+\caption{Energy consumption of an idle device with different radios enabled and applications running in background (\emph{Apps background} - Google Apps and Facebook for the test)}
 \label{tab:energy}
@@
 The different values are presented in Table.~\ref{tab:energy}\footnote{Energy consumption provided as average current drawn because it is independent of current battery voltage}.
-Energy consumption of an idle device with applications in background (\emph{Apps background}) considers a device with Google Apps, and Facebook installed and running in background and is consistent with our earlier observations~\cite{Aucinas:2013uk}.
-We have only run the test on a single WiFi network, therefore we expect the figure to vary, especially because our used network is a public one with some background traffic affecting energy consumption of the device.
 
-With the default implementation of Remote SIM Access over BlueTooth, the device consumes three times the amount of energy of an idle device connected to a 3G network, which is a high energy cost to pay, however the amount is comparable to that of an idle device with multiple applications running in background.
+The absolute values will vary with different devices and network configuration or conditions and application use, but the overall results remain.
+In particular, energy consumption of the different idle radios is similar, but cellular networks are significantly less efficient for current applications' network usage patterns~\cite{Aucinas:2013uk}.
+Therefore, replacing the cellular link with a WiFi one to another, powered device can bring significant savings - 32.5\% in our example case.
 
-Since we have implemented SIM sharing over WiFi using our custom client, we verified that the BlueTooth version is unnecessarily wasteful - once the PhoneLet has registered to the network all communication with the SIM server (for authentication) happens on demand, \eg when a new phonecall is received after 30 minutes of no activity.
-We therefore expect energy consumption of an optimized implementation over WiFi to be very close to that of an idle WiFi channel.
-For comparison, energy consumption of WiFi tethering even with a single idle client is significantly higher than that of sharing connectivity by offloading the phone functionality.
+The default BlueTooth implementation of RSAP is not efficient enough - maintaining access alone consumes almost three times the amount of energy of an idle device connected to a 3G network.
+Instead, we have implemented SIM sharing over WiFi using our custom client and verified that the BlueTooth version is unnecessarily wasteful - once the PhoneLet has registered to the network, there is no communication with the SIM server for as long as the device is idle.
+For example, a phonecall is successfully received after 30 minutes of no activity.
+Current Android platform does not allow RSAP to run over WiFi, but based on our measurements we expect energy consumption of an optimized implementation to be very close to that of an idle WiFi channel - 3.5x less than BlueTooth.
 
-These are early results, and we have not yet been able to replace the BlueTooth channel for an Android device.
-Nevertheless, the results point to a potential use-case where a powered \emph{PhoneLet} would be registered to the network on behalf of a WiFi-connected client with a considerable energy benefit.
-The benefit is further improved when network access is shared across multiple such devices and SIM access load is balanced across them.
-The approach is highly suitable for providing wide-area network connectivity for a larger number of devices where other options are not available.f
+Finally, since RSAP can be seen as an alternative to WiFi tethering when the mobile client provides connectivity to other devices, we measured energy consumption of WiFi tethering.
+With just one idle client it is almost 3x higher than the BlueTooth case.
+RSAP over WiFi effectively reverses the roles of client and server (\emph{PhoneLet} becomes the server) and similarly as above we expect energy consumption of an optimized WiFi mobile client to be close to 10x less than WiFi tethering.
+
+These are early results, but they point to a potential use-case where a powered \emph{PhoneLet} is registered to the network on behalf of a WiFi-connected client at a considerable energy benefit.
+Outside the scope of this demo, the benefit would be further improved when network access is shared across multiple such devices and SIM access load is balanced across them as the savings for each individual device would add up.
 
@@
 
-In the three steps we show our prototype's ability to offload phone functionality off a phone - share subscriber credentials between multiple clients and select which one will be connected to the network.
+In the three steps we show our prototype's ability to offload phone functionality from a phone - share subscriber credentials between multiple clients and select which one will be connected to the network.
 
@@
 
-If we only consider Internet connectivity, one could claim that \emph{PhoneLets} can be replaced by simple tethering as supported by all modern smartphones, however it would not take into account battery consumption, resource savings from sharing a single connection or even the lower energy consumption of sharing the SIM card than sharing Internet connectivity as discussed above.
+In addition to connectivity, \emph{PhoneLets} offers management of an individual's ID - the phone number.
+There have been multiple offerings with such function: multiple SIM cards associated with a number or VoIP-based solutions (\eg Google Voice), however they do not solve the problem of an increasing number of cellular clients and billing depends on service provider.
 
-\emph{PhoneLets} also offers management of part of an individual's ID - the phone number.
-In the past, multiple providers have offered a service in the past whereby multiple SIM cards would be associated with the number and multiple handsets would ring on an incoming call, \eg Google Voice, however it does not solve the problem of an increasing number of cellular clients and billing depends on the exact agreement with the service provider.
-One cellular network in UK has even offered a service to make phonecalls from tablets and laptops over WiFi to its users using a special app\footnote{O2 TU Go \url{http://www.o2.co.uk/apps/tu-go}}.
+One cellular network in UK has offered a service to make phonecalls from tablets and laptops over WiFi to its users using a special app\footnote{O2 TU Go \url{http://www.o2.co.uk/apps/tu-go}} further blurring the line between mobile and the Internet.
+If WiFi Internet connectivity is available, the approach could help reduce energy consumption by disabling cellular radio.
 
-Reprogrammable SIM cards~\cite{OFcom:2012tx} and more recently embedded SIM~\cite{Association:2013ub} have been proposed as a solution to the provisioning problem for M2M communication.
-Essentially, it still uses physical secure element to store credentials, but provides mechanisms to manage them over the air, primarily for long lifetime applications where replacing SIM cards is difficult.
-The architecture, unsurprisingly, allows operators to retain full control over the credentials, including when new ones can be installed or old ones removed, necessitating centralized management and reducing flexibility on the client side.
+Reprogrammable SIM cards~\cite{OFcom:2012tx} such as embedded SIM~\cite{Association:2013ub} have been proposed for provisioning in M2M communication.
+They maintain a physical secure element to store credentials, but provide mechanisms to manage them over the air.
+The architecture, allows operators to retain full control over the credentials, including when new ones can be installed or old ones removed, with necessarily centralized management and reduced flexibility on the client side.
 
@@
 
-A peculiar feature of the smartphone has been the decreasing importance of it as a phone, instead becoming an Internet device.
-Subscription to cellular networks, however, becomes expensive and energy-consuming for multiple devices.
-\emph{PhoneLets} is a system that allows for offloading of phone functionality of a cellular device, instead sharing a single subscription and therefore a single link amongst multiple devices in order to save costs and energy, decreasing cellular network infrastructure load at the same time.
+Increasingly, smartphones are replicating computer use, with phone now a very secondary function to email, social media and entertainment.
+However mobile network connections remain expensive both financially and in terms of the energy they use.
+\emph{PhoneLets} is a system that allows a single mobile subscription to be shared between multiple devices, allowing much of the mobile functionality of a cellular device to be offloaded.
+This allows the financial and energy costs of a cellular connection to be shared between multiple devices.
 
@@
 (2) \emph{Two standard UK power sockets}.
-(3) \emph{Two cellular network subscriptions (SIM cards)}, but are to be sources by the presenters, (4) WiFi AP for local communication.
+(3) \emph{Two cellular network subscriptions (SIM cards)}, which are to be sourced by the presenters, (4) WiFi AP for local communication.
 Setup time - up to an hour for connecting the demo up and verifying operation.

### 2014-07-01 paper.tex
submitted version

@@
 %make title bold and 14 pt font (Latex default is non-bold, 16 pt)
-\title{PhoneLets: offloading the phone off your phone for energy, cost and network load optimization}
+\title{Demo: PhoneLets -- offloading the phone off your phone for energy, cost and network load optimization}
 \ifnum\anon=1
@@
 Cellular connectivity has its own problems.
-This demo shows how some of them can be mitigated by offloading cellular connectivity to a \emph{PhoneLet} - sharing one SIM (Subscriber Identity Module) between nearby devices.
+This demo shows how PhoneLets allow one SIM (Subscriber Identity Module) to be shared between nearby devices.
+This addresses the following problems with cellular connectivity:
 
-Firstly, it is relatively expensive for the users~\cite{Anonymous:2013ut} and typically the end user needs to obtain a separate subscription for each of his devices connected to the cellular network, e.g.
+Firstly, connectivity is relatively expensive for users~\cite{Anonymous:2013ut}, typically requiring a separate subscription for each device connected to the cellular network, e.g.
 smartphone and tablet as well as broadband connectivity at home and at work.
-Each device often has its own subscription and fees and sharing a single SIM across them solves the issue.
+Using PhoneLets to share a single SIM across devices solves this issue.
 
-Secondly, the price is largely driven by the high infrastructure costs which increase with the growing number of network users - the increasing number of connected, periodically communicating clients also increases cellular network load, in the extreme cases causing \emph{signaling storms}~\cite{Yang:wh} that can bring down an entire network.
-Our approach reduces the number of simultaneously connected clients by sharing the same cellular link across all clients.
+Secondly, the increasing number of connected, periodically communicating clients increases infrastructure costs and also increases cellular network load, in extreme cases causing \emph{signaling storms}~\cite{Yang:wh} that can bring down an entire network.
+PhoneLets reduces the number of simultaneously connected clients by sharing the same cellular link across multiple clients.
 
-Finally, energy consumption of always connected clients is high compared to short-range radio technologies.
-Sharing a cellular link also helps reduce it for mobile devices.
-In previous work we analyzed the power consumption of cellular communication~\cite{Aucinas:2013uk} and showed that careful optimization is necessary to prolong the battery life of mobile devices.
-We discuss in Section~\ref{sec:energy} how PhoneLets can help with such optimization.
+Finally, energy consumption of always connected clients is high compared to short-range radio technologies. Sharing a cellular link also helps reduce it for mobile devices.
 
-\emph{PhoneLets} enable phone offloading by exploiting remote SIM (RSAP) functionality of existing hardware.
+\emph{PhoneLets} enable phone offloading by exploiting remote SIM functionality of existing hardware.
 A cellular module exchange commands with a SIM/UICC module over a wireless communication link rather than direct, physical connection.
-Some uses of such technology in the past have included automotive as well as electronic currency~\cite{subramanian2003sim}.
+Uses of such technology have included automotive as well as electronic currency~\cite{subramanian2003sim}.
 \emph{PhoneLets} infrastructure establishes and manages secure communication channels between participating entities and allows a device to connect to a cellular network without a physical SIM module being present.
@@
 
-Our design extends the standardized RSAP version by using a SIM server which is not capable of mobile connectivity itself and by using arbitrary communication links to access a remote SIM card (\emph{no SIM} in the mobile client).
+Our design extends the standardized RSAP (Remote SIM Access Protocol) version by using a SIM server which is not capable of mobile connectivity itself and by using arbitrary communication links to access a remote SIM card (\emph{no SIM} in the mobile client).
 Our work for this demo focused primarily on the client and SIM server design, and we are using off-the-shelf hardware as the mobile server to demonstrate backwards-compatibility.
@@
 \includegraphics[width=0.9\columnwidth]{figs/sequence}
-\vspace{-2mm}
+% \vspace{-2mm}
 \caption{Communication sequence between client, server and remote SIM card}
 \label{fig:sequence}
-\vspace{-4mm}
+% \vspace{-4mm}
 \end{figure}
 
-Mobile Client can obtain network credentials from any server -- a user's mobile phone, a co-located or a remote SIM server -- as long as it can reach it.
+The mobile Client can obtain network credentials from any server -- a user's mobile phone, a co-located or a remote SIM server -- as long as it can reach it.
 The key observation is that there are few operations that require access to the SIM card due to its interface speed (it would be infeasible to use it on the data path) and instead it is used to generate encryption keys.
@@
 \includegraphics[width=0.9\columnwidth]{figs/client}
-\vspace{-2mm}
+% \vspace{-2mm}
 \caption{Client prototype}
 \label{fig:client}
-\vspace{-3mm}
+% \vspace{-3mm}
 \end{figure}
@@
 
-Cost and network resource savings of using one active subscriber instead of multiple are straightforward, but reduced mobile client energy consumption is an important motivation for \emph{PhoneLets}.
+Reduced mobile client energy consumption is an important motivation for \emph{PhoneLets}.
 We analyze energy consumption of the client in a few example scenarios using our early prototype.
 The \emph{PhoneLet} itself is not yet energy-optimized, i.e.
-the controlling RaspberryPI is not meant for mobile uses, we focused on energy comparison of the mobile SIM server - a phone carried by the user from which we offload phone functionality.
+the controlling RaspberryPI is not meant for mobile uses, so we focused on energy comparison of the mobile SIM server - a phone carried by the user from which we offload phone functionality.
 
@@
 }
-\vspace{-2mm}
-\caption{Energy consumption of an idle device with different radios enabled and applications running in background (\emph{Apps background} - Google Apps and Facebook for the test)}
+% \vspace{-2mm}
+\caption{Energy consumption of an idle device with different radios enabled. (Apps background = Google apps and Facebook)}
 \label{tab:energy}
-\vspace{-3mm}
+% \vspace{-3mm}
 \end{table}
@@
 For example, a phonecall is successfully received after 30 minutes of no activity.
-Current Android platform does not allow RSAP to run over WiFi, but based on our measurements we expect energy consumption of an optimized implementation to be very close to that of an idle WiFi channel - 3.5x less than BlueTooth.
+The current Android platform does not allow RSAP to run over WiFi, but based on our measurements we expect energy consumption of an optimized implementation to be very close to that of an idle WiFi channel - 3.5x less than BlueTooth.
 
-Finally, since RSAP can be seen as an alternative to WiFi tethering when the mobile client provides connectivity to other devices, we measured energy consumption of WiFi tethering.
+Finally, we measured energy consumption of WiFi tethering.
 With just one idle client it is almost 3x higher than the BlueTooth case.
-RSAP over WiFi effectively reverses the roles of client and server (\emph{PhoneLet} becomes the server) and similarly as above we expect energy consumption of an optimized WiFi mobile client to be close to 10x less than WiFi tethering.
+RSAP is effectively an alternative to WiFi tethering when the mobile client provides connectivity to other devices.
+So similarly to above, we expect energy consumption of an optimized WiFi mobile client to be close to 10x less than WiFi tethering.
 

### 2014-07-13 paper.tex
camera-ready in progress

@@
 % This version uses the latex2e styles, not the very ancient 2.09 stuff.
-\documentclass{sig-alternate-10pt}
+% \documentclass{sig-alternate-10pt}
+\documentclass{sig-alternate-2013}
 \usepackage{endnotes,url}
@@
 \usepackage{listings}
+\DeclareCaptionType{copyrightbox}
 
@@
 %
-\ifnum\shownotes=1
+\ifnum\shownotes=0
 \newcommand{\authnote}[2]{{ $\ll$\textsf{\footnotesize #1 notes: #2}$\gg$}}
@@
 \usepackage{enumitem}
-\setlist{nolistsep}
+% \setlist{nolistsep}
 
 %Model parameters
+
+\clubpenalty=10000 
+\widowpenalty = 10000 
+\newfont{\mycrnotice}{ptmr8t at 7pt}
+\newfont{\myconfname}{ptmri8t at 7pt}
+\let\crnotice\mycrnotice%
+\let\confname\myconfname%
+\permission{Permission to make digital or hard copies of part or all of this work for personal or classroom use is granted without fee provided that copies are not made or distributed for profit or commercial advantage, and that copies bear this notice and the full citation on the first page.
+Copyrights for third-party components of this work must be honored.
+For all other uses, contact the owner/author(s).
+Copyright is held by the author/owner(s).}
+\conferenceinfo{MobiCom'14,}{September 7-11, 2014, Maui, Hawaii, USA.} 
+\copyrightetc{ACM \the\acmcopyr}
+\crdata{ACM 978-1-4503-2783-1/14/09. \\
+http://dx.doi.org/10.1145/2639108.2641744}
 
@@
 %make title bold and 14 pt font (Latex default is non-bold, 16 pt)
-\title{Demo: PhoneLets -- offloading the phone off your phone for energy, cost and network load optimization}
+\title{Demo: PhoneLets -- Offloading the Phone off your Phone for Energy, Cost and Network Load Optimization}
 \ifnum\anon=1
@@
 \alignauthor Andrius Aucinas\\
+\affaddr{Computer Laboratory}\\
 \affaddr{University of Cambridge}
+\email{andrius.aucinas@cl.cam.ac.uk} 
 \and
 \alignauthor Jon Crowcroft\\
+\affaddr{Computer Laboratory}\\
 \affaddr{University of Cambridge}
+\email{jon.crowcroft@cl.cam.ac.uk}
 }
@@
 
+\category{C.2.3}{Computer-communication networks}{Network Operations}{Network Management}
+
+\keywords{Mobile; Energy; Cellular; Networks; SIM; Subscriber Management}
+
 \section{What PhoneLets are}
@@
 Cellular connectivity has its own problems.
-This demo shows how PhoneLets allow one SIM (Subscriber Identity Module) to be shared between nearby devices.
+This demo shows how \emph{PhoneLets} allow one SIM (Subscriber Identity Module) to be shared between nearby devices.
 This addresses the following problems with cellular connectivity:
@@
 smartphone and tablet as well as broadband connectivity at home and at work.
-Using PhoneLets to share a single SIM across devices solves this issue.
+Using \emph{PhoneLets} to share a single SIM across devices solves this issue.
 
 Secondly, the increasing number of connected, periodically communicating clients increases infrastructure costs and also increases cellular network load, in extreme cases causing \emph{signaling storms}~\cite{Yang:wh} that can bring down an entire network.
-PhoneLets reduces the number of simultaneously connected clients by sharing the same cellular link across multiple clients.
+\emph{PhoneLets} reduce the number of simultaneously connected clients by sharing the same cellular link across multiple clients.
 
@@
 Uses of such technology have included automotive as well as electronic currency~\cite{subramanian2003sim}.
-\emph{PhoneLets} infrastructure establishes and manages secure communication channels between participating entities and allows a device to connect to a cellular network without a physical SIM module being present.
+\emph{PhoneLets} infrastructure establishes and manages secure communication channels between participating entities and allows a device to connect to a cellular network without a physical SIM card.
 
@@
 \includegraphics[width=0.9\columnwidth]{figs/arch}
-\vspace{-3mm}
-\caption{Architecture}
+\caption{Architecture of PhoneLets and interaction with different types of servers}
 \label{fig:arch}
-\vspace{-2mm}
 \end{figure}
@@
 
-Our design extends the standardized RSAP (Remote SIM Access Protocol) version by using a SIM server which is not capable of mobile connectivity itself and by using arbitrary communication links to access a remote SIM card (\emph{no SIM} in the mobile client).
+The SIM card is a special I/O device in a mobile platform: due to its security requirements it is not exposed to the platform's operating system.
+Instead, it is managed by the modem's firmware which routes all communication between the network and the subscriber module.
+Standard remote IO approaches are similar in their nature (\eg RIO~\cite{Sani:2014hs} enables transparently accessing I/O devices of a remote mobile system) but are not applicable in this setting.
+
+Interaction with the subscriber module requires lower level protocols understood by the modem firmware.
+Our design extends the standard RSAP (Remote SIM Access Protocol)~\cite{Anonymous:2008vz} by using a SIM server which is not capable of mobile connectivity itself and by using arbitrary communication links to access a remote SIM card (\emph{no SIM} in the mobile client).
 Our work for this demo focused primarily on the client and SIM server design, and we are using off-the-shelf hardware as the mobile server to demonstrate backwards-compatibility.
@@
 \centering
-\includegraphics[width=0.9\columnwidth]{figs/sequence}
-% \vspace{-2mm}
+\includegraphics[width=0.95\columnwidth]{figs/sequence}
 \caption{Communication sequence between client, server and remote SIM card}
 \label{fig:sequence}
-% \vspace{-4mm}
 \end{figure}
 
-The mobile Client can obtain network credentials from any server -- a user's mobile phone, a co-located or a remote SIM server -- as long as it can reach it.
+The mobile client can obtain network credentials from any server -- a user's mobile phone, a co-located or a remote SIM server -- as long as it can reach it.
 The key observation is that there are few operations that require access to the SIM card due to its interface speed (it would be infeasible to use it on the data path) and instead it is used to generate encryption keys.
@@
 \centering
-\includegraphics[width=0.9\columnwidth]{figs/client}
-% \vspace{-2mm}
+\includegraphics[width=0.95\columnwidth]{figs/client}
 \caption{Client prototype}
 \label{fig:client}
-% \vspace{-3mm}
 \end{figure}
 
-Our demo mobile client (Fig.~\ref{fig:client}), the \emph{PhoneLet}, consists of a Telit UC864 cellular module connected to a raspberryPI as a controller, which also uses WiFi and BlueTooth dongles for local wireless connectivity.
+Our demo mobile client (Fig.~\ref{fig:client}), the \emph{PhoneLet}, consists of a Telit UC864 cellular module connected to a raspberryPI as a controller, which also uses WiFi and BlueTooth (BT) dongles for local wireless connectivity.
 The controller runs a modified version of the Ofono open telephony stack and manages communication between the cellular module and multiple backends.
 
-The two servers are an off-the-shelf Samsung Galaxy S4 and our custom server that only interfaces with SIM cards through a smartcard reader, also running on a raspberryPI.
+The two example servers are an off-the-shelf Samsung Galaxy S4 and our custom server that only interfaces with SIM cards through a smartcard reader, also running on a raspberryPI.
 In our prototype, the server is connected to a WiFi AP to be reachable by the mobile client.
-We replace the BlueTooth communication channel with a more secure option - TCP link with certificate-based authentication of both client and server, and strong encryption for all data exchange.
+We replace the BT communication channel with a more secure option - TCP link with certificate-based authentication of both client and server, and strong encryption for all data exchange.
 Once the client establishes a connection to either backend, it registers with the network and takes on the role of a phone.
@@
 }
-% \vspace{-2mm}
 \caption{Energy consumption of an idle device with different radios enabled. (Apps background = Google apps and Facebook)}
 \label{tab:energy}
-% \vspace{-3mm}
 \end{table}
@@
 Energy consumption of a Samsung Galaxy S4 was measured\footnote{Using Monsoon Power Monitor \url{http://www.msoon.com/LabEquipment/PowerMonitor/}} in a number of scenarios.
-The different values are presented in Table.~\ref{tab:energy}\footnote{Energy consumption provided as average current drawn because it is independent of current battery voltage}.
+The different values are presented in Table.~\ref{tab:energy}\footnote{Current drain (mA) is chosen over power (mW) as power depends on voltage that decays with the battery discharge}.
 
@@
 
-The default BlueTooth implementation of RSAP is not efficient enough - maintaining access alone consumes almost three times the amount of energy of an idle device connected to a 3G network.
-Instead, we have implemented SIM sharing over WiFi using our custom client and verified that the BlueTooth version is unnecessarily wasteful - once the PhoneLet has registered to the network, there is no communication with the SIM server for as long as the device is idle.
+The default BT implementation of RSAP is not efficient enough - maintaining access alone consumes almost three times the amount of energy of an idle device connected to a 3G network.
+Instead, we have implemented SIM sharing over WiFi using our custom client and verified that the BT version is unnecessarily wasteful - once the PhoneLet has registered to the network, there is no communication with the SIM server for as long as the device is idle.
 For example, a phonecall is successfully received after 30 minutes of no activity.
-The current Android platform does not allow RSAP to run over WiFi, but based on our measurements we expect energy consumption of an optimized implementation to be very close to that of an idle WiFi channel - 3.5x less than BlueTooth.
+The current Android platform does not allow RSAP to run over WiFi, but based on our measurements we expect energy consumption of an optimized implementation to be very close to that of an idle WiFi channel - 3.5x less than BT.
 
 Finally, we measured energy consumption of WiFi tethering.
-With just one idle client it is almost 3x higher than the BlueTooth case.
+With just one idle client it is almost 3x higher than the BT case.
 RSAP is effectively an alternative to WiFi tethering when the mobile client provides connectivity to other devices.
@@
 
+It is well known that WiFi tethering consumes a lot of energy.
+Cool-Tether~\cite{Sharma:2009fd} proposed to reverse the roles of a mobile gateway and client and turn the client into WiFi AP.
+Our proposal also reduces energy consumption of the mobile gateway by completely changing the roles of the entities.
+Instead, the client (\emph{PhoneLet}) uses the cellular network directly, avoiding additional wireless communication with the gateway.
+
 Reprogrammable SIM cards~\cite{OFcom:2012tx} such as embedded SIM~\cite{Association:2013ub} have been proposed for provisioning in M2M communication.
@@
 
-%We highly encourage that each proposal includes a video clip showcasing the demo.
-The video clip, no more than 60 seconds, should give a good idea of what the demo is about and how it would look like.
-If you do include a video clip, it will help the committee better understand and evaluate your proposal.
-If your demo is accepted, the video will likely be shown on the conference website and during the conference to draw people to your booth.
-Videos can be submitted by including a link to the video (e.g., hosted at YouTube).
-
-\section*{Resource Requirements}
-
-(1) \emph{Desk/table of approximately 2m, standard-width} to accommodate the demo.
-(2) \emph{Two standard UK power sockets}.
-(3) \emph{Two cellular network subscriptions (SIM cards)}, which are to be sourced by the presenters, (4) WiFi AP for local communication.
-Setup time - up to an hour for connecting the demo up and verifying operation.
-Other equipment includes a laptop, two smartphones and the custom-build hardware described above.
-
-% Equipment to be used for the demo.
-% Space needed and setup time required.
-% Additional facilities needed including power and Internet/wireless access.
-
 %Only show things we really cite :)
 % \nocite{*}
-{\footnotesize 
+{ 
 \bibliographystyle{abbrv}

### 2014-07-15 paper.tex
hopefully camera-ready

@@
 \usepackage{enumitem}
-% \setlist{nolistsep}
+\setlist{nolistsep}
 
@@
 
-\clubpenalty=10000 
-\widowpenalty = 10000 
 \newfont{\mycrnotice}{ptmr8t at 7pt}
@@
 \copyrightetc{ACM \the\acmcopyr}
-\crdata{ACM 978-1-4503-2783-1/14/09. \\
+\crdata{978-1-4503-2783-1/14/09. \\
 http://dx.doi.org/10.1145/2639108.2641744}
+\clubpenalty=10000 
+\widowpenalty = 10000
 
@@
 This demo presents how phone functionality can be offloaded from a smartphone over wireless link to a \emph{PhoneLet} by sharing one SIM card across multiple devices.
-This can lead to significant cost and network load reductions by reducing the number of simultaneously connected mobile clients.
-Furthermore, it can bring significant energy reductions for the mobile user when connected to a powered \emph{PhoneLet}.
-It absorbs the energy cost of inefficient mobile applications' communication patterns, instead providing connectivity for the user over a WiFi link.
+This can lead to significant cost and network load reductions by decreasing the number of simultaneously connected mobile clients.
+Furthermore, it can save energy for the mobile user when connected to a powered \emph{PhoneLet} by offloading phone functionality.
+It absorbs the energy cost of online presence and inefficient mobile applications' communication patterns, instead providing connectivity for the user over a WiFi link.
 \end{abstract}
@@
 \centering
-\includegraphics[width=0.9\columnwidth]{figs/arch}
+\includegraphics[width=\columnwidth]{figs/arch}
 \caption{Architecture of PhoneLets and interaction with different types of servers}
@@
 
-The overall concept and system architecture is shown in Fig.~\ref{fig:arch}.
+Figure~\ref{fig:arch} shows the overall concept and system architecture.
 There are three types of entities: 1) mobile server, a mobile device which can share its SIM with other clients, 2) SIM server which may host and provide more sets of credentials over local or wide-area networks and 3) a mobile client which can use either of the methods.
@@
 Instead, it is managed by the modem's firmware which routes all communication between the network and the subscriber module.
-Standard remote IO approaches are similar in their nature (\eg RIO~\cite{Sani:2014hs} enables transparently accessing I/O devices of a remote mobile system) but are not applicable in this setting.
+Standard remote IO approaches are similar in their nature (\eg RIO~\cite{Sani:2014hs} enables transparently accessing I/O devices of a remote mobile system) but are not applicable in this setting as they require OS support.
 
@@
 Our design extends the standard RSAP (Remote SIM Access Protocol)~\cite{Anonymous:2008vz} by using a SIM server which is not capable of mobile connectivity itself and by using arbitrary communication links to access a remote SIM card (\emph{no SIM} in the mobile client).
-Our work for this demo focused primarily on the client and SIM server design, and we are using off-the-shelf hardware as the mobile server to demonstrate backwards-compatibility.
+Our work for this demo focused primarily on the client and SIM server design.
+We use off-the-shelf hardware as the mobile server for backwards-compatibility.
 
@@
 \centering
-\includegraphics[width=0.95\columnwidth]{figs/sequence}
+\includegraphics[width=\columnwidth]{figs/sequence}
 \caption{Communication sequence between client, server and remote SIM card}
@@
 \centering
-\includegraphics[width=0.95\columnwidth]{figs/client}
+\includegraphics[width=\columnwidth]{figs/client}
 \caption{Client prototype}
@@
 
-Energy consumption of a Samsung Galaxy S4 was measured\footnote{Using Monsoon Power Monitor \url{http://www.msoon.com/LabEquipment/PowerMonitor/}} in a number of scenarios.
-The different values are presented in Table.~\ref{tab:energy}\footnote{Current drain (mA) is chosen over power (mW) as power depends on voltage that decays with the battery discharge}.
+We measured energy consumption of a Samsung Galaxy S4\footnote{Using Monsoon Power Monitor \url{http://www.msoon.com/LabEquipment/PowerMonitor/}} in a number of scenarios.
+Table.~\ref{tab:energy}\footnote{Current drain (mA) is chosen over power (mW) as power depends on voltage that decays with the battery discharge} presents energy consumption of the different device radios, depending on their use, ranging from an idle device with only a particular radio enabled, to different methods of sharing connectivity and simple application usage patterns.
 
@@
 Instead, we have implemented SIM sharing over WiFi using our custom client and verified that the BT version is unnecessarily wasteful - once the PhoneLet has registered to the network, there is no communication with the SIM server for as long as the device is idle.
-For example, a phonecall is successfully received after 30 minutes of no activity.
-The current Android platform does not allow RSAP to run over WiFi, but based on our measurements we expect energy consumption of an optimized implementation to be very close to that of an idle WiFi channel - 3.5x less than BT.
+For example, a phone call is successfully received after 30 minutes of no activity.
+The current Android platform does not allow RSAP to run over WiFi, but based on our measurements we expect energy consumption of an optimized implementation to be close to that of an idle WiFi channel - 3.5x less than BT.
 
@@
 RSAP is effectively an alternative to WiFi tethering when the mobile client provides connectivity to other devices.
-So similarly to above, we expect energy consumption of an optimized WiFi mobile client to be close to 10x less than WiFi tethering.
+Therefore we expect energy consumption of an optimized WiFi mobile client to be close to 10x less than WiFi tethering.
 
 These are early results, but they point to a potential use-case where a powered \emph{PhoneLet} is registered to the network on behalf of a WiFi-connected client at a considerable energy benefit.
-Outside the scope of this demo, the benefit would be further improved when network access is shared across multiple such devices and SIM access load is balanced across them as the savings for each individual device would add up.
+The benefit would be further improved when network access is shared across multiple such devices and SIM access load is balanced across them as the savings for each individual device would add up.
+This use-case is outside the scope of the current demo.
 
@@
 
-One cellular network in UK has offered a service to make phonecalls from tablets and laptops over WiFi to its users using a special app\footnote{O2 TU Go \url{http://www.o2.co.uk/apps/tu-go}} further blurring the line between mobile and the Internet.
+One cellular network in UK has offered a service to make phonecalls from tablets and laptops over WiFi to its users using a special application\footnote{O2 TU Go \url{http://www.o2.co.uk/apps/tu-go}}, blurring the line between mobile and the Internet.
 If WiFi Internet connectivity is available, the approach could help reduce energy consumption by disabling cellular radio.
@@
 They maintain a physical secure element to store credentials, but provide mechanisms to manage them over the air.
-The architecture, allows operators to retain full control over the credentials, including when new ones can be installed or old ones removed, with necessarily centralized management and reduced flexibility on the client side.
+The architecture allows operators to retain full control over the credentials, including when new ones can be installed or old ones removed through centralized management and subsequently reduced flexibility on the client side.
 
